---
title: "spline_script"
author: "yacine"
date: "2024-10-11"
output: html_document
---

We load and/or install the packages we will be using:

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Refresh environment 


rm(list= ls())
#install.packages("splines2")
#install.packages("rms")
#install.packages('Greg')
library(dplyr)
library(survival)
library(readr)
library(ggplot2)
library(splines)
library(splines2)
library(rms)
library(Greg)
library(data.table)
library(readr)
library(dvmisc)
library(gtsummary)
library(gt)
library(kableExtra)
library(broom.helpers)
library(flextable)
library(officer)
library(ggplot2)
library(splines)
library(splines2)
library(survminer)

splines::bs
splines2::bSpline

#theme_gtsummary_journal("jama")
```

## We load the data:
```{r}
#dt <- readRDS("/home/yacine/UKBB_beluga/prepped_dt.rds") # full df with exclusions + recoding and all data prepared for survival analysis
dt <- readRDS("/home/yacine/dataframes_AccelerometerProcessing/prepped_dt.rds")
dt2 <- readRDS("/home/yacine/dataframes_AccelerometerProcessing/prepped_dt2.rds")
dt3 <- readRDS("/home/yacine/dataframes_AccelerometerProcessing/prepped_dt3.rds")
```


## Spline Cox PH fit Using the rms Package:
```{r}

# Renaming the variables for MVPA and Follow-up time (crucial for coxph function to work)
dt <- dt %>% rename(fu_time = `Follow up time in months`,
                    MI = `Myocardial infarction`,
                    mvpa1 = `MVPA min/week - ENMO`,
                    mvpa2 = `MVPA min/week - Machine learning`,
                    mvpa3 = `MVPA min/week - Activity count`,
                    deprivation_index = `Deprivation index`,
                    hh_income = `Household income GBP`,
                    Alcohol = `Alcohol consumption`,
                    pro_meat = `Processed meat`,
                    red_meat = `Red meat`,
                    fresh_fruit = `Fresh fruit`,
                    cooked_vg = `Cooked vegetables`,
                    oily_fish = `Oily fish`,
                    non_oily_fish = `Non-oily fish`,
                    greenspace = `Greenspace 300m from residence`,
                    water_res = `Water 300m from residence`,
                    natural_env = `Natural env. 300m from residence`
                    #type_2_diab = `Type II Diabetes`,
                    #bmi = `Body Mass Index`
                    )

dt2 <- dt2 %>% rename(fu_time = `Follow up time in months`,
                    MI = `Myocardial infarction`,
                    mvpa1 = `MVPA min/week - ENMO`,
                    mvpa2 = `MVPA min/week - Machine learning`,
                    mvpa3 = `MVPA min/week - Activity count`,
                    mvpa4 = `MVPA min/week - Self-report (IPAQ)`,
                    deprivation_index = `Deprivation index`,
                    hh_income = `Household income GBP`,
                    Alcohol = `Alcohol consumption`,
                    pro_meat = `Processed meat`,
                    red_meat = `Red meat`,
                    fresh_fruit = `Fresh fruit`,
                    cooked_vg = `Cooked vegetables`,
                    oily_fish = `Oily fish`,
                    non_oily_fish = `Non-oily fish`,
                    #type_2_diab = `Type II diabetes`,
                    #bmi = `Body mass index`,
                    greenspace = `Greenspace 300m from residence`,
                    water_res = `Water 300m from residence`,
                    natural_env = `Natural env. 300m from residence`
)

dt3 <- dt3 %>% rename(fu_time = `Follow up time in months`,
                    MI = `Myocardial infarction`,
                    mvpa1 = `MVPA min/week - ENMO`,
                    mvpa2 = `MVPA min/week - Machine learning`,
                    mvpa3 = `MVPA min/week - Activity count`,
                    deprivation_index = `Deprivation index`,
                    #hh_income = `Household income GBP`,
                    #pro_meat = `Processed meat`,
                    #red_meat = `Red meat`,
                    #fresh_fruit = `Fresh fruit`,
                    #cooked_vg = `Cooked vegetables`,
                    #oily_fish = `Oily fish`,
                    #non_oily_fish = `Non-oily fish`,
                    type_2_diab = `Type II diabetes`,
                    bmi = `Body mass index`,
                    greenspace = `Greenspace 300m from residence`,
                    water_res = `Water 300m from residence`,
                    natural_env = `Natural env. 300m from residence`
                    )

# *** Note: We determined the optimal number of knots (nk) for each PA measure using AIC testing.The function only allows nk >= 3. Therefore, the number of knots used corresponds to the lowest AIC value for each.
# As a result, we determined the optimal nk was 4 for PA1 and PA2, and 3 for PA3 and PA4.
```

## Main analysis

### Stroke and MI
Does not include diabetes, bmi and depression and PA4
```{r}

# PA1
## AIC(nk=3) > AIC(nk=4) 
fit_pa1_stroke <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + red_meat + greenspace + water_res + natural_env  , data = dt)


# PA2
## AIC(nk=3) > AIC(nk=4) 
fit_pa2_stroke <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt)

#fit_pa2_stroke_nk4 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + red_meat, data = dt)


#PA3
## AIC(nk=3) < AIC(nk=4) 
fit_pa3_stroke <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa3, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt)


```


### MI
```{r}

# PA1
#AIC(nk=3) > AIC(nk=4)

fit_pa1_MI <- coxph(Surv(fu_time, MI) ~ rcs(mvpa1, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt)



# PA2
#AIC(nk=3) < AIC(nk=4)
fit_pa2_MI <- coxph(Surv(fu_time, MI) ~ rcs(mvpa2, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt)



#PA3
#AIC(nk=3) < AIC(nk=4)
fit_pa3_MI <- coxph(Surv(fu_time, MI) ~ rcs(mvpa3, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt)


```

Function to label the plots: 
```{r}

# source: https://logfc.wordpress.com/2017/03/15/adding-figure-labels-a-b-c-in-the-top-left-corner-of-the-plotting-region/

fig_label <- function(text, region = "plot", pos = "topleft", cex = 1.2, x_adj = -0.1, y_adj = -0.03, ...) {
  region <- match.arg(region, c("figure", "plot", "device"))
  pos <- match.arg(pos, c("topleft", "top", "topright",
                          "left", "center", "right",
                          "bottomleft", "bottom", "bottomright"))
  
  if (region == "plot") {
    u <- par("usr")  
    x <- u[1:2]
    y <- u[3:4]
  } else {
    stop("This function is configured for 'plot' region only.")
  }
  

  x1 <- switch(pos,
               topleft = x[1] + (x[2] - x[1]) * x_adj,
               left = x[1] + (x[2] - x[1]) * x_adj,
               bottomleft = x[1] + (x[2] - x[1]) * x_adj,
               top = (x[1] + x[2]) / 2,
               center = (x[1] + x[2]) / 2,
               bottom = (x[1] + x[2]) / 2,
               topright = x[2] - (x[2] - x[1]) * x_adj,
               right = x[2] - (x[2] - x[1]) * x_adj,
               bottomright = x[2] - (x[2] - x[1]) * x_adj)
  
  y1 <- switch(pos,
               topleft = y[2] + (y[1] - y[2]) * y_adj,
               top = y[2] + (y[1] - y[2]) * y_adj,
               topright = y[2] + (y[1] - y[2]) * y_adj,
               left = (y[1] + y[2]) / 2,
               center = (y[1] + y[2]) / 2,
               right = (y[1] + y[2]) / 2,
               bottomleft = y[1] - (y[1] - y[2]) * y_adj,
               bottom = y[1] - (y[1] - y[2]) * y_adj,
               bottomright = y[1] - (y[1] - y[2]) * y_adj)
  
  old.par <- par(xpd = NA)
  on.exit(par(old.par))
  
  text(x1, y1, text, cex = cex, font = 2, ...)
  return(invisible(c(x, y)))
}

```




#### Combined Plots - Main Analysis
```{r}
png("combined_main.png", width = 2250, height = 2625, res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(3, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))

# PA1
## Stroke
plotHR(fit_pa1_stroke,  
       term = 1, 
       xlab = "MVMVPA duration (minutes/week)", 
       adj = 0, 
       xlim = c(0, 1500), 
       ylim = c(0.2, 4), 
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa1_MI,
       term = 1, 
       xlab = "MVMVPA duration (minutes/week)", 
       xlim = c(0, 1500), 
       ylim = c(0.2, 4), 
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


# PA2
## Stroke
plotHR(fit_pa2_stroke, 
       term = 1, 
       xlab = "MVPA puration (minutes/week)", 
       xlim = c(0, 1500), 
       ylim = c(0.2, 4), 
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa2_MI, 
       term = 1, 
       xlab = "MVMVPA duration (minutes/week)", 
       xlim = c(0, 1500), 
       ylim = c(0.2, 4), 
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


# PA3 
## Stroke
plotHR(fit_pa3_stroke, 
       term = 1, 
       xlab = "MVMVPA duration (minutes/week)", 
       xlim = c(0, 1500), 
       ylim = c(0.2, 4), 
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("E", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


## MI
plotHR(fit_pa3_MI, term = 1, 
       xlab = "MVMVPA duration (minutes/week)", 
       xlim = c(0, 1500), 
       ylim = c(0.2, 4), 
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("F", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

dev.off()

```



* Sensitivity analysis:

We create formulas for covariate (they are currently not working in the rcs function, but ultimately I would like to add them to the functions)
```{r}

f_cov <- "Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + type_2_diab + bmi + Depression"

f_cov_noMediator <- "Sex + Age + Race + deprivation_index + Education + hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish"

#as.formula(paste("Surv(`Follow_up time`,`Myocardial Infarction`) ~ ", "MVPA_Quant_PA2 + ",  f_cov))  
#as.formula(paste(f_stroke, " ~" ,  "MVPA_Quant_PA2 + ",  f_cov))

```



## Sensitivity analysis 1 - No Mediator, without BMI and Diabetes and Depression

### Stroke
```{r}

# PA1
fit_pa1_stroke_sa1 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat + non_oily_fish + greenspace + water_res + natural_env , data = dt2)

#fit_pa1_stroke_sa1 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat + non_oily_fish + greenspace + water_res + natural_env , data = dt2)

# PA2
fit_pa2_stroke_sa1<- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish + greenspace + water_res + natural_env , data = dt2)

#fit_pa2_stroke_nk4_sa1 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish, data = dt)


#PA3
fit_pa3_stroke_sa1 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa3, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish + greenspace + water_res + natural_env , data = dt2)


#PA4 
fit_pa4_stroke_sa1 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa4, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish + greenspace + water_res + natural_env , data = dt2)



```


### MI

```{r}
# PA1
fit_pa1_mi_sa1 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa1, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish +red_meat + non_oily_fish + greenspace + water_res + natural_env , data = dt2)

#fit_pa1_stroke_sa1 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat + non_oily_fish + greenspace + water_res + natural_env , data = dt2)

# PA2
fit_pa2_mi_sa1 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa2, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish + greenspace + water_res + natural_env , data = dt2)

#fit_pa2_mi_nk4_sa1 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa2, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish +red_meat + non_oily_fish, data = dt)


#PA3
fit_pa3_mi_sa1 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa3, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish +red_meat + non_oily_fish + greenspace + water_res + natural_env , data = dt2)


#PA4 
fit_pa4_mi_sa1 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa4, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish + greenspace + water_res + natural_env , data = dt2)


```

#### Combined Plots

8 plots in one figure
```{r}
png("combined_sa1.png", width = 2250, height = 2625, res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(4, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))

# PA1
plotHR(fit_pa1_stroke_sa1,
       term = 1,
       xlab = "MVMVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA1 MI
plotHR(fit_pa1_mi_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA2
plotHR(fit_pa2_stroke_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA2 MI
plotHR(fit_pa2_mi_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA3
## Stroke
plotHR(fit_pa3_stroke_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, cex.lab = 1, cex.main = 1)
fig_label("E", pos = "topleft", cex = 1.5, x_adj = -0.05, y_adj = -0.03)

## MI
plotHR(fit_pa3_mi_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, cex.lab = 1, cex.main = 1, family = "Arial"  # Added font control
)
fig_label("F", pos = "topleft", cex = 1.5, x_adj = -0.05, y_adj = -0.03)

# PA4
## Stroke
plotHR(fit_pa4_stroke_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 2000),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, cex.lab = 1, cex.main = 1, family = "Arial"  # Added font control
)
fig_label("G", pos = "topleft", cex = 1.5, x_adj = -0.05, y_adj = -0.03)

## MI
plotHR(fit_pa4_mi_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 2000),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, cex.lab = 1, cex.main = 1, family = "Arial"  # Added font control
)
fig_label("H", pos = "topleft", cex = 1.5, x_adj = -0.05, y_adj = -0.03)

dev.off()


```

2 figures of four plots
```{r}
png("combined_sa1_1.png", width = 2200, height = 2250,res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(2, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))

# PA1
plotHR(fit_pa1_stroke_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA1 MI
plotHR(fit_pa1_mi_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA2
plotHR(fit_pa2_stroke_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA2 MI
plotHR(fit_pa2_mi_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

dev.off()


png("combined_sa1_2.png", width = 2200, height = 2250, res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(2, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))

# PA3
## Stroke
plotHR(fit_pa3_stroke_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, cex.lab = 1, cex.main = 1)
fig_label("E", pos = "topleft", cex = 1.5, x_adj = -0.05, y_adj = -0.03)

## MI
plotHR(fit_pa3_mi_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, cex.lab = 1, cex.main = 1, family = "Arial"  # Added font control
)
fig_label("F", pos = "topleft", cex = 1.5, x_adj = -0.05, y_adj = -0.03)

# PA4
## Stroke
plotHR(fit_pa4_stroke_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 2000),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, cex.lab = 1, cex.main = 1, family = "Arial"  # Added font control
)
fig_label("G", pos = "topleft", cex = 1.5, x_adj = -0.05, y_adj = -0.03)

## MI
plotHR(fit_pa4_mi_sa1,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 2000),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, cex.lab = 1, cex.main = 1, family = "Arial"  # Added font control
)
fig_label("H", pos = "topleft", cex = 1.5, x_adj = -0.05, y_adj = -0.03)

dev.off()

```


## Sensitivity analysis 2 - Removed people with huge duration (>1200min) of MVPA Stroke 
### Stroke
```{r}

dt_trunc_1200 <- dt %>%  
  mutate( MVPA_min_PA1_truncated_1200 = ifelse(MVPA_min_PA1_truncated > 1200, NA, MVPA_min_PA1_truncated), 
          MVPA_min_PA2_truncated_1200 = ifelse(MVPA_min_PA2_truncated > 1200, NA, MVPA_min_PA2_truncated),
          MVPA_min_PA3_truncated_1200 = ifelse(MVPA_min_PA3_truncated > 1200, NA, MVPA_min_PA3_truncated)
          ) 


dt_trunc_1200 <- dt_trunc_1200[!is.na(dt_trunc_1200$MVPA_min_PA1_truncated_1200), ]
dt_trunc_1200 <- dt_trunc_1200[!is.na(dt_trunc_1200$MVPA_min_PA2_truncated_1200), ]
dt_trunc_1200 <- dt_trunc_1200[!is.na(dt_trunc_1200$MVPA_min_PA3_truncated_1200), ]



# PA1
fit_pa1_stroke_sa2 <- coxph(Surv(fu_time, Stroke) ~ rcs(MVPA_min_PA1_truncated_1200, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish +red_meat + non_oily_fish + greenspace + water_res + natural_env , data = dt_trunc_1200)


# PA2
fit_pa2_stroke_sa2 <- coxph(Surv(fu_time, Stroke) ~ rcs(MVPA_min_PA2_truncated_1200, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish +red_meat + non_oily_fish + greenspace + water_res + natural_env , data = dt_trunc_1200)

#fit_pa2_stroke_nk4_sa2 <- coxph(Surv(fu_time, Stroke) ~ rcs(MVPA_min_PA2_truncated_1200, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish +red_meat + non_oily_fish, data = dt2)


#PA3
fit_pa3_stroke_sa2 <- coxph(Surv(fu_time, Stroke) ~ rcs(MVPA_min_PA3_truncated_1200, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt_trunc_1200)


```
### MI
```{r}
# PA1
fit_pa1_MI_sa2 <- coxph(Surv(fu_time, MI) ~ rcs(MVPA_min_PA1_truncated_1200, 4) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt_trunc_1200)


# PA2
fit_pa2_MI_sa2 <- coxph(Surv(fu_time, MI) ~ rcs(MVPA_min_PA2_truncated_1200, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + red_meat + greenspace + water_res + natural_env , data = dt_trunc_1200)



#PA3
fit_pa3_MI_sa2 <- coxph(Surv(fu_time, MI) ~ rcs(MVPA_min_PA3_truncated_1200, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt_trunc_1200)

```
#### Combined Plots
```{r}
png("combined_sa2.png", width = 2250, height = 2625, res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(3, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))

# PA1
## Stroke
plotHR(fit_pa1_stroke_sa2,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
      cex.axis = 1, cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa1_MI_sa2,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA2
plotHR(fit_pa2_stroke_sa2,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
      cex.axis = 1, cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa2_MI_sa2,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

# PA3
## Stroke
plotHR(fit_pa3_stroke_sa2,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("E", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa3_MI_sa2,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("F", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

dev.off()



```



## Sensitivity analysis 3 - Male and Female separate 
See the data source in the regression command

### Female
```{r}
# PA1
fit_pa1_stroke_sa3 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 4) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Female", ])

fit_pa1_MI_sa3 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa1, 4) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Female", ])

# PA2
fit_pa2_stroke_sa3 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Female", ])

fit_pa2_MI_sa3 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa2, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Female", ])

# PA3
fit_pa3_stroke_sa3 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa3, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Female", ])

fit_pa3_MI_sa3 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa3, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Female", ])


```
### Male
```{r}
# PA1
fit_pa1_stroke_sa3_m <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Male", ])

#fit_pa1_stroke_sa3_m <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 4) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Male", ])

fit_pa1_MI_sa3_m <- coxph(Surv(fu_time, MI) ~ rcs(mvpa1, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Male", ])

#fit_pa1_MI_sa3_m <- coxph(Surv(fu_time, MI) ~ rcs(mvpa1, 4) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Male", ])

# PA2
fit_pa2_stroke_sa3_m <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Male", ])

fit_pa2_MI_sa3_m <- coxph(Surv(fu_time, MI) ~ rcs(mvpa2, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Male", ])

# PA3
fit_pa3_stroke_sa3_m <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa3, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Male", ])

fit_pa3_MI_sa3_m <- coxph(Surv(fu_time, MI) ~ rcs(mvpa3, 3) + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish +red_meat + greenspace + water_res + natural_env , data = dt[dt$Sex == "Male", ])


```

#### Combined Plots

```{r}

# PA1
png("combined_sa3_pa1.png", width = 2200, height = 2250, res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(2, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))
## Stroke - Female
plotHR(fit_pa1_stroke_sa3,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


## MI - Female
plotHR(fit_pa1_MI_sa3,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


## Stroke - Male
plotHR(fit_pa1_stroke_sa3_m,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)



## MI - Male
plotHR(fit_pa1_MI_sa3_m,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)
dev.off()



#######################################
# PA2
png("combined_sa3_pa2.png", width = 2200, height = 2250,res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(2, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))
## Stroke - Female
plotHR(fit_pa2_stroke_sa3,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)



## MI - Female
plotHR(fit_pa2_MI_sa3,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U" ,
       cex.axis = 1, cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


## Stroke - Male
plotHR(fit_pa2_stroke_sa3_m,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


## MI
plotHR(fit_pa2_MI_sa3_m,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

dev.off()



#######################################
# PA3
png("combined_sa3_pa3.png", width = 2200, height = 2250, res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(2, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))
## Stroke - Female
plotHR(fit_pa3_stroke_sa3,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


## MI - Female
plotHR(fit_pa3_MI_sa3,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",,
       cex.axis = 1, cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## Stroke - Male
plotHR(fit_pa3_stroke_sa3_m,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI - Male
plotHR(fit_pa3_MI_sa3_m,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U",
       cex.axis = 1, cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)
dev.off()


```


## Sensitivity analysis 4 - 2 years of survival 
We perform a sensitivity analysis by first extracting people that have had a stroke or MI
within the 2 following years of the end of the accelerometer tracking ('2_years_survival_stroke' == 0;
or, 2_years_survival_MI' == 0)

Remove people who have had a Stroke or MI within 2 years following date_end_accel:
```{r}

# we create a new df where we extracted the participants for whom '2_years_survival_stroke' == 0
dt_sa4 <- dt %>%
  filter(`2_years_survival_stroke` != 0)

dt_sa4 <- dt_sa4  %>%
  filter(`2_years_survival_MI` != 0)

```

### Stroke
```{r}
# PA1
fit_pa1_stroke_sa4 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + greenspace + water_res + natural_env , data = dt_sa4 )

# PA2
fit_pa2_stroke_sa4 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + greenspace + water_res + natural_env , data = dt_sa4 )

# PA3
fit_pa3_stroke_sa4 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa3, 3) + Sex + Age + Race +  deprivation_index + Education +  hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + greenspace + water_res + natural_env , data = dt_sa4 )



```


### MI
```{r}
# PA1
fit_pa1_MI_sa4 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa1, 3) + Sex + Age + Race + deprivation_index + Education + hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + greenspace + water_res + natural_env , data = dt_sa4)

# PA2
fit_pa2_MI_sa4 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa2, 3) + Sex + Age + Race + deprivation_index + Education + hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + greenspace + water_res + natural_env , data = dt_sa4)

# PA3
fit_pa3_MI_sa4 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa3, 3) + Sex + Age + Race + deprivation_index + Education + hh_income + Smoking + Alcohol + pro_meat + fresh_fruit + cooked_vg + oily_fish + non_oily_fish + greenspace + water_res + natural_env , data = dt_sa4)


```
#### Combined Plots
```{r}
png("combined_sa4.png",width = 2250, height = 2625, res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(3, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))

# PA1
## Stroke
plotHR(fit_pa1_stroke_sa4,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa1_MI_sa4,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


# PA2
plotHR(fit_pa2_stroke_sa4,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa2_MI_sa4,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


# PA3
## Stroke
plotHR(fit_pa3_stroke_sa4,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("E", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa3_MI_sa4,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("F", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

dev.off()


```


## Sensitivity analysis 5 - Forrest Plots

In the "3_survival_analysis_prep.Rmd" file



## Sensitivity analysis 6 - Including Mediator (w/o PA4)

### Stroke
```{r}
# PA1
fit_pa1_stroke_sa6 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa1, 4) + Sex + Age + ethnicity +  deprivation_index + education_level +  avg_hh_income + smoking + alcohol + processed_meat + fresh_fruit + cooked_vg + oily_fish + red_meat + non_oily_fish + greenspace + water_res + natural_env  + bmi + type_2_diab + Depression, data = dt3)


# PA2
fit_pa2_stroke_sa6<- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 3) + Sex + Age + ethnicity +  deprivation_index + education_level +  avg_hh_income + smoking + alcohol + processed_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish + greenspace + water_res + natural_env  + bmi + type_2_diab + Depression, data = dt3)

#fit_pa2_stroke_nk4_sa1 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa2, 4) + Sex + Age + ethnicity +  deprivation_index + Education +  hh_income + Smoking + Alcohol +processed_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish, data = dt)


#PA3
fit_pa3_stroke_sa6 <- coxph(Surv(fu_time, Stroke) ~ rcs(mvpa3, 3) + Sex + Age + ethnicity +  deprivation_index + education_level +  avg_hh_income + smoking + alcohol + processed_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish + greenspace + water_res + natural_env  + bmi + type_2_diab + Depression, data = dt3)

```

### MI
```{r}
# PA1
fit_pa1_MI_sa6 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa1, 4) + Sex + Age + ethnicity +  deprivation_index + education_level +  avg_hh_income + smoking + alcohol + processed_meat + fresh_fruit + cooked_vg + oily_fish + red_meat + non_oily_fish + greenspace + water_res + natural_env  + bmi + type_2_diab + Depression, data = dt3)


# PA2
fit_pa2_MI_sa6<- coxph(Surv(fu_time, MI) ~ rcs(mvpa2, 3) + Sex + Age + ethnicity +  deprivation_index + education_level +  avg_hh_income + smoking + alcohol + processed_meat + fresh_fruit + cooked_vg + oily_fish + red_meat +non_oily_fish + greenspace + water_res + natural_env  + bmi + type_2_diab + Depression, data = dt3)


#PA3
fit_pa3_MI_sa6 <- coxph(Surv(fu_time, MI) ~ rcs(mvpa3, 3) + Sex + Age + ethnicity +  deprivation_index + education_level +  avg_hh_income + smoking + alcohol + processed_meat + fresh_fruit + cooked_vg + oily_fish + greenspace + water_res + natural_env  + red_meat +non_oily_fish + bmi + type_2_diab + Depression, data = dt3)

```

#### Combined Plots
```{r}
png("combined_sa6.png", width = 2250, height = 2625, res = 300)
par(family = "Arial", ps = 12, cex = 1, mfrow = c(3, 2), oma = c(0, 0, 2, 0), mar = c(4, 4, 2, 1))

# PA1
## Stroke
plotHR(fit_pa1_stroke_sa6,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("A", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


## MI
plotHR(fit_pa1_MI_sa6,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("B", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


# PA2
plotHR(fit_pa2_stroke_sa6,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("C", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa2_MI_sa6,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("D", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)


# PA3
## Stroke
plotHR(fit_pa3_stroke_sa6,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("E", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

## MI
plotHR(fit_pa3_MI_sa6,
       term = 1,
       xlab = "MVPA duration (minutes/week)",
       xlim = c(0, 1500),
       ylim = c(0.2, 4),
       plot.bty = "U", 
       cex.axis = 1, 
       cex.lab = 1.2)
fig_label("F", pos = "topleft", cex = 1.5, x_adj = -0.1, y_adj = -0.03)

dev.off()


```
