---
title: "survival_analysis_prep"
author: "yacine"
date: "2024-07-16"
output:
  pdf_document: default
  html_document: default
---

We load and/or install the packages we will be using:

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Refresh environment 
rm(list= ls())
library(dplyr)
library(survival)
library(data.table)
library(readr)
library(dvmisc)
library(gtsummary)
library(gt)
library(kableExtra)
library(broom.helpers)
library(flextable)
library(officer)
library(ggplot2)
library(splines)
library(splines2)
library(survminer)
#install.packages("splines2")
splines::bs
splines2::bSpline

theme_gtsummary_journal("jama")
```

## We load the data:
```{r}
#dt <- readRDS("/home/yacine/UKBB_beluga/prepped_dt.rds") # full df with exclusions + recoding and all data prepared for survival analysis
dt <- readRDS("/home/yacine/dataframes_AccelerometerProcessing/prepped_dt.rds")
dt2 <- readRDS("/home/yacine/dataframes_AccelerometerProcessing/prepped_dt2.rds")
```



###################
## Forest Plot Function 
We create the forest plot function which we will be using to visualize the regression analysis 

```{r}


funcPlotForest <- function(forest_data, title = "Forest Plot", subtitle = "Hazard Ratios and 95% Confidence Intervals", maxX = 2) {

  forest_data <- forest_data %>%
    filter(!is.na(est) & !is.na(lwr) & !is.na(upr)) %>%
    mutate(
      index = factor(index, levels = rev(unique(index))),
      foodName = factor(foodName, levels = c( "Vector Magnitude (ENMO)","Random Forest and HMM (ML)", "Activity Count"))
    ) 
  
  xname <- "Hazard Ratio and 95% CI"
  
  ggplot(data = forest_data, aes(y = index, x = est, xmin = lwr, xmax = upr, color = foodName)) + 
    geom_point(size = 3) + 
    geom_errorbarh(height = 0.4) +
    scale_x_continuous(limits = c(0.1, maxX), breaks = seq(0.1, maxX, by = 0.2), name = xname) +
    scale_y_discrete(name = NULL) +  # Ensure no title on y-axis
    geom_vline(xintercept = 1, color = "black", linetype = "dashed", alpha = 0.5) +
    scale_color_manual(values = c("Vector Magnitude (ENMO)" = "blue", "Random Forest and HMM (ML)" = "red", "Activity Count" = "brown")) +  # Set colors for groups
    theme_classic() + 
    theme(
axis.text.x = element_text(color = "black", size = 14, angle = 0, hjust = 0.5, vjust = 0.5, face = "plain"),
      axis.text.y = element_text(color = "black", size = 14, angle = 0, hjust = 1, vjust = 0, face = "plain"),
      axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),
      axis.title.y = element_text(color = "black", size = 14, angle = 0, hjust = 0.5, vjust = 0.5, face = "plain"),
      plot.title = element_text(color = "black", size = 20, face = "bold"),
      plot.subtitle = element_text(color = "black", size = 20),
      legend.text = element_text(size = 20),
      legend.title = element_blank() 
    ) +
    labs(title = title, subtitle = subtitle) +
    theme(
      strip.text.y = element_text(size = 20, vjust = 1),
      strip.background = element_rect(colour = "white", fill = "white")
    ) +
    theme(strip.text.y.left = element_text(angle = 0)) +
    theme(strip.placement = "outside") +
    coord_cartesian(xlim = c(0.1, maxX))  
}


```
  

Function forest plot 2
```{r}
funcPlotForest2 <- function(forest_data, maxX = 2, new_titles = NULL, tag_label = "A") { #remove title and subtitle to respect PLOS requirements
  

  forest_data <- forest_data %>%
    filter(!is.na(est) & !is.na(lwr) & !is.na(upr))
  

  if (!is.null(new_titles)) {
    forest_data$index <- new_titles
  }
  

  forest_data <- forest_data %>%
    mutate(
      index = factor(index, levels = rev(unique(index))),
      foodName = factor(foodName, levels = c("ENMO", "Machine learning", "Activity Count", "Self-Report (IPAQ)"))
    )
  

  xname <- "Hazard Ratio and 95% CI"
  

  ggplot(data = forest_data, aes(y = index, x = est, xmin = lwr, xmax = upr, color = foodName)) + 
    geom_point(size = 3) + 
    geom_errorbarh(height = 0.4) +
    scale_x_continuous(limits = c(0.1, maxX), breaks = seq(0.1, maxX, by = 0.2), name = xname) +
    scale_y_discrete(name = NULL) +  # no title for y-axis (PLOS requirements)
    geom_vline(xintercept = 1, color = "black", linetype = "dashed", alpha = 0.5) +
    scale_color_manual(values = c(
      "ENMO" = "blue",
      "Machine learning" = "red",
      "Activity Count" = "brown",
      "Self-Report (IPAQ)" = "black"
    )) +  
    theme_classic() + 
    theme( # size 12 text for PLOS requirements
      axis.text.x = element_text(color = "black", size = 12, angle = 0, hjust = 0.5, vjust = 0.5, face = "plain"),
      axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),
      axis.title.x = element_text(color = "black", size = 12, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),
      axis.title.y = element_text(color = "black", size = 12, angle = 0, hjust = 0.5, vjust = 0.5, face = "plain"),
     # plot.title = element_text(color = "black", size = 20, face = "bold"),
     # plot.subtitle = element_text(color = "black", size = 20),
      legend.text = element_text(size = 12),
      legend.title = element_blank(),
      plot.tag = element_text(face = "bold", size = 20) # letter labels font
    ) +
    labs(tag = tag_label) +  # add letter labels
    coord_cartesian(xlim = c(0.1, maxX))  
}

new_titles <- c(
  "ENMO - Q2",
  "ENMO - Q3",
  "ENMO- Q4",
  "Machine learning - Q2",
  "Machine learning - Q3",
  "Machine learning - Q4",
  "Activity Count - Q2",
  "Activity Count - Q3",
  "Activity Count - Q4"
)

```
  
  
  
## Regression - Associations with risk of incident cardiovascular disease 

In the data preparation step, we added an event status indicator at exit and a follow-up time variable. Using these, we can run a Cox model to associate overall activity with risk of incident cardiovascular disease. We'll start by using time-on-study as the timescale and set it up using the 'survival' package in R. We'll also adjust for various possible confounding variables (following the confounders used by [Ramakrishnan et al.](https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.1003487)):


First , create formula 
```{r}

f_cov <- "Sex + Age + Race + `Deprivation index` + Education + `Household income GBP` + Smoking + `Alcohol consumption` + `Processed meat` + `Fresh fruit` + `Cooked vegetables` + `Oily fish` + `Non-oily fish`"

f_cov_noMediator <- "Sex + Age + Race + `Deprivation index` + Education + `Household income GBP` + Smoking + `Alcohol consumption` + `Processed meat` + `Fresh fruit` + `Cooked vegetables` + `Oily fish` + `Non-oily fish`"

f_stroke <-   "Surv(`Follow up time in months`,`Myocardial infarction`)"

f_MI <- "Surv(`Follow up time in months`,`Stroke`)"


#as.formula(paste("Surv(`Follow_up time`,`Myocardial Infarction`) ~ ", "MVPA_Quant_PA2 + ",  f_cov))  
#as.formula(paste(f_stroke, " ~" ,  "MVPA_Quant_PA2 + ",  f_cov))



```
  

### Stroke - Main
```{r}

# Model for PA 1 
fitDf1 <- coxph(as.formula(paste(f_stroke, " ~" ,  "MVPA_Quant_PA1 + ",  f_cov)), data = dt)
tb1 <- tbl_regression(fitDf1, exponentiate = TRUE, 
                     label = list("MVPA_Quant_PA1" = "MVPA Quartile"))

# Model for PA 2
fitDf2 <- coxph(as.formula(paste(f_stroke, " ~" ,  "MVPA_Quant_PA2 + ",  f_cov)), data = dt)
tb2 <- tbl_regression(fitDf2, exponentiate = TRUE, 
                     label = list("MVPA_Quant_PA2" = "MVPA Quartile"))

# Model for PA 3
fitDf3 <- coxph(as.formula(paste(f_stroke, " ~" ,  "MVPA_Quant_PA3 + ",  f_cov)), data = dt)
tb3<- tbl_regression(fitDf3, exponentiate = TRUE, 
                     label = list("MVPA_Quant_PA3" = "MVPA Quartile"))


#Merge tables 
tb_merge <- tbl_merge(
  tbls = list(tb1, tb2, tb3),
  tab_spanner = c("**Vector Magnitude (ENMO)**", "**Random Forest and HMM**", "**Activity Count**")) %>% 
  bold_labels() %>% 
   modify_column_hide(columns = c(p.value_1, p.value_2, p.value_3))


# Landscape (word format)
tb_main_stroke <- as_flex_table(tb_merge)
sect_properties <- prop_section(
  page_size = page_size(
  orient = "portrait",
  #width = 8, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar(0) 
  )

```
#### Forest Plot 
```{r}
# we first need to convert data from tbl_resgression to data.frame to be able 
# to use the generated "funcPlotForest" function and we combine the two
data_t1 <- as.data.frame(tb1$table_body)
data_t2 <- as.data.frame(tb2$table_body)
data_t3 <- as.data.frame(tb3$table_body)


data_t1 <- data_t1 %>%
  rename(
    index = term,
    est = estimate,
    lwr = conf.low,
    upr = conf.high
  ) %>%
  mutate(
    foodName = "ENMO"  
  )

data_t2 <- data_t2 %>%
  rename(
    index = term,
    est = estimate,
    lwr = conf.low,
    upr = conf.high
  ) %>%
  mutate(
    foodName = "Machine learning" 
  )

data_t3 <- data_t3 %>%
  rename(
    index = term,
    est = estimate,
    lwr = conf.low,
    upr = conf.high
  ) %>%
  mutate(
    foodName = "Activity Count" 
  )



forest_data <- bind_rows(data_t1, data_t2, data_t3)
forest_data <- forest_data %>%
  filter(variable %in% c("MVPA_Quant_PA1", "MVPA_Quant_PA2", "MVPA_Quant_PA3"))


# we can now create the plot using the funcPlotForest function created earlier:
png("rf_sensa5_stroke.png", width = 2250, height = 2625, res = 300)
par(family = "Arial", ps = 12)

funcPlotForest2(
  forest_data = forest_data,
  maxX = 1.3, 
  new_titles = new_titles,
  tag_label = "A" # specify label of figure
)

dev.off()
# Save the plot
#ggsave("/home/yacine/AcclerometerProcessing/survival_analysis/plots/sensa_1_stroke.png", plot, width = 10, height = 6, dpi = 300)

```



### MI Main 
```{r}

# Model for PA 1 
fitDf1 <- coxph(as.formula(paste(f_MI, " ~" ,  "MVPA_Quant_PA1 + ",  f_cov)), data = dt)
tb1 <- tbl_regression(fitDf1, exponentiate = TRUE, 
                        label = list("MVPA_Quant_PA1" = "MVPA Quartile"))

# Model for PA 2
fitDf2 <- coxph(as.formula(paste(f_MI, " ~" ,  "MVPA_Quant_PA2 + ",  f_cov)), data = dt)
tb2 <- tbl_regression(fitDf2, exponentiate = TRUE, 
                        label = list("MVPA_Quant_PA2" = "MVPA Quartile"))

# Model for PA 3
fitDf3 <- coxph(as.formula(paste(f_MI, " ~" ,  "MVPA_Quant_PA3 + ",  f_cov)), data = dt)
tb3 <- tbl_regression(fitDf3, exponentiate = TRUE, 
                        label = list("MVPA_Quant_PA3" = "MVPA Quartile"))



```
#### Forest Plot 
```{r}
# we first need to convert data from tbl_resgression to data.frame to be able 
# to use the generated "funcPlotForest" function and we combine the two
data_t1 <- as.data.frame(tb1$table_body)
data_t2 <- as.data.frame(tb2$table_body)
data_t3 <- as.data.frame(tb3$table_body)


data_t1 <- data_t1 %>%
  rename(
    index = term,
    est = estimate,
    lwr = conf.low,
    upr = conf.high
  ) %>%
  mutate(
    foodName = "ENMO"  # or another identifier if applicable
  )

data_t2 <- data_t2 %>%
  rename(
    index = term,
    est = estimate,
    lwr = conf.low,
    upr = conf.high
  ) %>%
  mutate(
    foodName = "Machine learning" 
  )

data_t3 <- data_t3 %>%
  rename(
    index = term,
    est = estimate,
    lwr = conf.low,
    upr = conf.high
  ) %>%
  mutate(
    foodName = "Activity Count" 
  )


forest_data <- bind_rows(data_t1, data_t2, data_t3)
forest_data <- forest_data %>%
  filter(variable %in% c("MVPA_Quant_PA1", "MVPA_Quant_PA2", "MVPA_Quant_PA3"))


# we can now create the plot using the funcPlotForest function created earlier:
png("rf_sensa5_MI.png", width = 2250, height = 2625, res = 300)
par(family = "Arial", ps = 12)

funcPlotForest2(
  forest_data = forest_data,
  maxX = 1.3, 
  new_titles = new_titles,
  tag_label = "B" # specify label of figure
)

dev.off()

# Save the plot
#ggsave("/home/yacine/AcclerometerProcessing/survival_analysis/plots/sensa_1_stroke.png", plot, width = 10, height = 6, dpi = 300)

```



