---
title: "survival_analysis_prep"
author: "yacine"
date: "2024-07-16"
output: rmarkdown::github_document
---

We load and/or install the packages we will be using:

```{r, include=FALSE}
library(dplyr)
```

We load the data:
```{r}
#full_df <- read.csv("/home/yacine/final_df_full.csv") # full df w/o exclusions
full_df <- read.csv("/home/yacine/UKBB_beluga/df_exc.csv") # full df with exclusions
```

We change the date format for the start and end accelerometer variables - which
will be crucial for later steps:
```{r}
full_df$date_start_accel <- as.Date(full_df$date_start_accel)
full_df$date_end_accel <- as.Date(full_df$date_end_accel)
```


## Follow-up time:

First we create the "end_of_fu" variable, which can be classified as date of 
stroke/MI, date of death or the end of the study (i.e. 2022):

```{r}
# we create a "end_of_study" column which indicates the end of the whole data collection:
full_df$end_of_study_date <- as.Date('2022-12-31')

# we reate a vector of all the date columns:
date_cols <- c('stroke_date', 'ischaemic_stroke_date', 'intracerebral_haemorrhage_date',
               'myocardial_infarction_date', 'STEMI_date', 'NSTEMI_date',
               'date_of_death_0', 'date_of_death_1', 'date_lost_followup',
               'date_end_accel','end_of_study_date')



# we make sure all values within the date columns are in date format:
full_df[date_cols] <- lapply(full_df[date_cols], as.Date)

# we create the end_of_fu column by only keeping the earliest date from each rows
# within each date columns:
full_df$end_of_fu <- apply(full_df[date_cols], 1, function(x) min(x[!is.na(x)]))

#print(full_df$end_of_fu)
```

We can now create the "follow-up time" column by calculating the difference (in months) 
betweenthe "end of follow up" and "start time of wear" columns:
```{r}
# remove NA data from date_start_accel (to do in data exclusion UKBB script?)
full_df <- full_df[!is.na(full_df$date_start_accel), ]

difftime_weeks <- difftime(full_df$end_of_fu, full_df$date_start_accel, units = "weeks")

# because the "difftime" function only allows a maximum of units in weeks, we 
# devide by 4.345 to convert to difference months
full_df$fu_time <- as.numeric(difftime_weeks) / 4.345

#head(full_df$fu_time) 
#print(full_df$date_start_accel)
```


##### Create binary variable for CVD (strokes and MI)

```{r}
# First we create a new column which will indicate if a participant has had: 
## a) a stroke == 1; or not == 0:
full_df$stroke <- ifelse(
  final_df$stroke_date == "" | is.na(final_df$ischaemic_stroke_date) | is.na(final_df$intracerebral_haemorrhage_date),
  0,
  1
)

## b) a MI == 1; or not == 0:
full_df$MI <- ifelse(
  final_df$myocardial_infarction_date == "NA" | is.na(final_df$STEMI_date) | is.na(final_df$NSTEMI_date),
  0,
  1
)

print(full_df$stroke)
print(full_df$MI)
```




################ GitHub PA code

# Now link data with acclerometer weekly minutes, for LPA, MVPA, and sedendary behavior.   
- There are 3 diferent accelerometric measures   
  1. PA1 - Vector magnitude based measures as calcuated from time-series of vector magnitude by UKBB 
  2. PA2 - machine-learning derived mesared, for Field ID 1020
  3. PA3 (pending) - Actigraph count-based measure, based on Freedson cutoff 


#Load derived accelerometer from UKBB, PA2 data 
```{r, include=FALSE}
ML_derived_PA2 <- fread("~/projects/UKBB/PA2_category_1020.csv")
ML_derived_PA2 <- ML_derived_PA2 %>%  dplyr::select(
  eid, 
  "40048-0.0",  
  "40049-0.0",
  "40047-0.0", 
  "40046-0.0", 
  "40044-0.0",  
  "40045-0.0",
  "40043-0.0", 
  "40042-0.0" 
)

# relabel for id, weekly light PA, weekly MVPA, weekly sleep, weekly sedentary, daily light PA, daily MVPA, daily sedentary
names(ML_derived_PA2) <- c("eid", "overall_l", "overall_mv", "overall_sb", "overall_slp", "day_l", "day_mv", "day_sb", "day_slp")

ML_derived_PA2 <- ML_derived_PA2 %>% filter(!is.na(overall_l))
dim(ML_derived_PA2)
```




# Files from parallel processing of time-series (FieldID 90004) to generate ENMO summary of 100,000 ppl, PA1 data 
```{r, include=FALSE}
library(readr)
enmoTSParallel_PA1 <- list.files(path="PA1_parallel_TS_ENMO/outputs/", full.names = TRUE) %>% 
  lapply(read_csv) %>% 
  bind_rows 
```










