---
title: "survival_analysis_prep"
author: "yacine"
date: "2024-07-16"
output: rmarkdown::github_document
---

We load and/or install the packages we will be using:

```{r, include=FALSE}

#Refresh environment 
rm(list= ls())

library(dplyr)
library(survival)
library(data.table)
library(readr)
library(dvmisc)

#library(survminer)
```

We load the data:
```{r}
#full_df <- read.csv("/home/yacine/final_df_full.csv") # full df w/o exclusions
#full_df <- read.csv("/home/yacine/UKBB_beluga/df_exc.csv") # full df with exclusions
full_df <- read.csv("/home/yacine/UKBB_beluga/df_exc_recode.csv") # full df with exclusions + recoding
```

We change the date format for the start and end accelerometer variables - which
will be crucial for later steps:
```{r}
full_df$date_start_accel <- as.Date(full_df$date_start_accel)
full_df$date_end_accel <- as.Date(full_df$date_end_accel)
```


## Follow-up time:

First we create the "end_of_fu" variable, which can be classified as date of 
stroke/MI, date of death or the end of the study (i.e. 2022):

```{r}
# we find out which death date is the latest which we will set as the "end of the study" 
# date
max_death_date_0 <- max(full_df$'date_of_death_0', na.rm = TRUE)
max_death_date_1 <- max(full_df$'date_of_death_1', na.rm = TRUE)
print(max_death_date_0)
print(max_death_date_1)

# given that we see that the max(death_date) is "2022-12-17", this will serve as 
# our "end of study" date

# we create a "end_of_study" column which indicates the end of the whole data collection:
full_df$end_of_study_date <- as.Date('2022-12-17')

# we reate a vector of all the date columns:
date_cols <- c('stroke_date', 'ischaemic_stroke_date', 'intracerebral_haemorrhage_date',
               'myocardial_infarction_date', 'STEMI_date', 'NSTEMI_date',
               'date_of_death_0', 'date_of_death_1', 'date_lost_followup',
               'end_of_study_date')



# we make sure all values within the date columns are in date format:
full_df[date_cols] <- lapply(full_df[date_cols], as.Date)

# we create the end_of_fu column by only keeping the earliest date from each rows
# within each date columns:
full_df$end_of_fu <- apply(full_df[date_cols], 1, function(x) min(x[!is.na(x)]))

#print(full_df$end_of_fu)
```

We can now create the "follow-up time" column by calculating the difference (in months) 
between the "end of follow up" and "end time of wear" columns:
```{r}
# remove NA data from date_end_accel (to do in data exclusion UKBB script?)
full_df <- full_df[!is.na(full_df$date_end_accel), ]

difftime_weeks <- difftime(full_df$end_of_fu, full_df$date_end_accel, units = "weeks")

# because the "difftime" function only allows a maximum of units in weeks, we 
# devide by 4.345 to convert to difference months
full_df$fu_time <- as.numeric(difftime_weeks) / 4.345

head(full_df$fu_time) 
#print(full_df$date_start_accel)
```


##### Create binary variable for CVD (strokes and MI)

```{r}
# First we create a new column which will indicate if a participant has had: 
## a) a stroke == 1; or not == 0:
stroke_cols <- c('stroke_date', 'ischaemic_stroke_date', 'intracerebral_haemorrhage_date')

# Create the 'stroke' column
full_df$stroke <- ifelse(rowSums(!is.na(full_df[, stroke_cols ])) > 0, 1, 0)


## b) a MI == 1; or not == 0:
MI_cols <- c('myocardial_infarction_date', 'STEMI_date', 'NSTEMI_date')
full_df$MI <- ifelse(rowSums(!is.na(full_df[, MI_cols ])) > 0, 1, 0)


head(full_df$stroke)
head(full_df$MI)

# We now create a "CVD" column which indicates if the participants has had a
# stroke and/or MI (==1) or not (==0)
full_df <- full_df %>%
  mutate(CVD = ifelse(stroke == 1 | MI == 1, 1, 0))
```


## Associations with risk of incident cardiovascular disease 

In the data preparation step, we added an event status indicator at exit and a follow-up time variable. Using these, we can run a Cox model to associate overall activity with risk of incident cardiovascular disease. We'll start by using time-on-study as the timescale and set it up using the 'survival' package in R. We'll also adjust for various possible confounding variables (following the confounders used by [Ramakrishnan et al.](https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.1003487)):
```{r}

#cox_model <- coxph(
  #Surv(fu_time, stroke, MI) ~ overall_activity_quarters + age_entry_years + sex + ethnicity + tdi_quarters + age_education + smoking + alcohol,
 # data = full_df
#)
#summary(cox_model)
```




################ GitHub PA code

# Now link data with acclerometer weekly minutes, for LPA, MVPA, and sedendary behavior.   
- There are 3 diferent accelerometric measures   
  1. PA1 - Vector magnitude based measures as calcuated from time-series of vector magnitude by UKBB 
  2. PA2 - machine-learning derived mesared, for Field ID 1020
  3. PA3 (pending) - Actigraph count-based measure, based on Freedson cutoff 


#Load derived accelerometer from UKBB, PA2 data 
```{r, include=FALSE}
ML_derived_PA2 <- fread("/home/yacine/survival_analysis/PA2_category_1020.csv")
ML_derived_PA2 <- ML_derived_PA2 %>%  dplyr::select(
  eid, 
  "40048-0.0",  
  "40049-0.0",
  "40047-0.0", 
  "40046-0.0", 
  "40044-0.0",  
  "40045-0.0",
  "40043-0.0", 
  "40042-0.0" 
)

# relabel for id, weekly light PA, weekly MVPA, weekly sleep, weekly sedentary, daily light PA, daily MVPA, daily sedentary
names(ML_derived_PA2) <- c("eid", "overall_l", "overall_mv", "overall_sb", "overall_slp", "day_l", "day_mv", "day_sb", "day_slp")

ML_derived_PA2 <- ML_derived_PA2 %>% filter(!is.na(overall_l))
dim(ML_derived_PA2)


ML_derived_PA2$MVPA_min <- ML_derived_PA2$overall_mv * 7 * 24 * 60 
ML_derived_PA2$MVPA_Quant <- dvmisc::quant_groups(ML_derived_PA2$MVPA_min, 4)

```




# Files from parallel processing of time-series (FieldID 90004) to generate ENMO summary of 100,000 ppl, PA1 data 
```{r, include=FALSE, message=FALSE}
enmoTSParallel_PA1 <- list.files(path="/home/yacine/survival_analysis/PA1_parallel_TS_ENMO/outputs", full.names = TRUE) %>% 
  lapply(read_csv) %>% 
  bind_rows 

enmoTSParallel_PA1$MVPA_min <- enmoTSParallel_PA1$MVPA_total* 7*24*60
enmoTSParallel_PA1$MVPA_Quant <- dvmisc::quant_groups(enmoTSParallel_PA1$MVPA_min, 4)

```


We merge the full_df and PA1 and PA2 df by eids
```{r}
full_df_PA1 <- inner_join(full_df, enmoTSParallel_PA1, by = "eid")
full_df_incl_PA1_PA2 <- inner_join(full_df_PA1, ML_derived_PA2, by = "eid")

names(full_df_incl_PA1_PA2) <- gsub("\\.x$", "_PA1", names(full_df_incl_PA1_PA2))
names(full_df_incl_PA1_PA2) <- gsub("\\.y$", "_PA2", names(full_df_incl_PA1_PA2))


write.csv(full_df_incl_PA1_PA2, "/home/yacine/survival_analysis/full_df_with_PA.csv", row.names = FALSE)
```




####### Create Summary of Baseline Characteristics Table ######
```{r}

##### data recoding

# recode the BMI variable 
full_df_incl_PA1_PA2 <- full_df_incl_PA1_PA2 %>%
  mutate(BMI = case_when(
    BMI < 18.5 ~ "Underweight (< 18.5 kg/m2)",
    BMI >= 18.5 & BMI < 25 ~ "Normal (18.5 kg/m2 to < 25 kg/m2)",
    BMI >= 25 & BMI < 30 ~ "Overweight (25 kg/m2 to < 30 kg/m2)",
    BMI >= 30 ~ "Obesity Class I, II or III (> 30 kg/m2)",
    TRUE ~ NA_character_  
  ))

# recode binary variables 
full_df_incl_PA1_PA2 <- full_df_incl_PA1_PA2 %>%
  mutate(sex = recode(sex, `1` = "Man", `0` = "Woman"))


# we create the summary baseline characteristics table
summ_base_char <- full_df_incl_PA1_PA2 %>%
  tbl_summary(
    include = c(sex, fresh_fruit, oily_fish, cooked_vg,
                processed_meat_0, alcohol, smoking, ethnicity, age_education_raw, stroke, MI, 
                diabetes_0, BMI, MVPA_min_PA1, MVPA_min_PA2),
    statistic = list(all_continuous() ~ "{mean} ± {sd}")
  ) %>%
  bold_labels()


print(summ_base_char)

```
##############################################################









MVPA quantiles:
We make a function to cut by quantile:
```{r}
qtile_cut <-  function(x, probs = seq(0, 1, 0.25), na.rm = TRUE, labels = NULL) {
    breaks <- quantile(x = x, probs = probs, na.rm = na.rm)
    out <- cut(x = x, breaks = breaks, labels = labels, right = FALSE, include.lowest = TRUE)
    return(out)
}
```

```{r}
#full_df$overall_activity_quarters <- qtile_cut(full_df$overall_activity, labels = c("Quarter 1", "Quarter 2", "Quarter 3", "Quarter 4"))
```


##########
Kaplan_Meier:
  - 2 groups: PA1 and PA2 or Stroke and MI?
  - d_stroke == 1; d_MI == 1
  - end_of_fu = censoring time
  
```{r}

km_stroke <- coxph(
  Surv(fu_time, stroke) ~ overall_activity_quarters + age_entry_years + sex + ethnicity + tdi_quarters + age_education + smoking + alcohol,
  data = full_df
)

```
  




