---
title: "survival_analysis_prep"
author: "yacine"
date: "2024-07-16"
output:
  pdf_document: default
  html_document: default
---

We load and/or install the packages we will be using:

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Refresh environment 
rm(list= ls())
library(dplyr)
library(survival)
library(data.table)
library(readr)
library(dvmisc)
library(gtsummary)
library(gt)
library(kableExtra)
#remotes::install_github("larmarange/broom.helpers")
library(broom.helpers)
library(flextable)
library(officer)

```

We load the data:
```{r}
#full_df <- read.csv("/home/yacine/final_df_full.csv") # full df w/o exclusions
#full_df <- read.csv("/home/yacine/UKBB_beluga/df_exc.csv") # full df with exclusions
full_df <- read.csv("/home/yacine/UKBB_beluga/df_exc_recode.csv") # full df with exclusions + recoding
```

We change the date format for the start and end accelerometer variables - which
will be crucial for later steps:
```{r}
full_df$date_start_accel <- as.Date(full_df$date_start_accel)
full_df$date_end_accel <- as.Date(full_df$date_end_accel)
```


## Follow-up time:

First we create the "end_of_fu" variable, which can be classified as date of 
stroke/MI, date of death or the end of the study (i.e. 2022):

```{r}
# we find out which death date is the latest which we will set as the "end of the study" 
# date
max_death_date_0 <- max(full_df$'date_of_death_0', na.rm = TRUE)
max_death_date_1 <- max(full_df$'date_of_death_1', na.rm = TRUE)
print(max_death_date_0)
print(max_death_date_1)

# given that we see that the max(death_date) is "2022-12-17", this will serve as 
# our "end of study" date

# we create a "end_of_study" column which indicates the end of the whole data collection:
full_df$end_of_study_date <- as.Date('2022-12-17')

# we reate a vector of all the date columns:
date_cols <- c('stroke_date', 'ischaemic_stroke_date', 'intracerebral_haemorrhage_date',
               'myocardial_infarction_date', 'STEMI_date', 'NSTEMI_date',
               'date_of_death_0', 'date_of_death_1', 'date_lost_followup',
               'end_of_study_date')



# we make sure all values within the date columns are in date format:
full_df[date_cols] <- lapply(full_df[date_cols], as.Date)

# we create the end_of_fu column by only keeping the earliest date from each rows
# within each date columns:
full_df$end_of_fu <- apply(full_df[date_cols], 1, function(x) min(x[!is.na(x)]))

#print(full_df$end_of_fu)
```

We can now create the "follow-up time" column by calculating the difference (in months) 
between the "end of follow up" and "end time of wear" columns:
```{r}
# remove NA data from date_end_accel (to do in data exclusion UKBB script?)
full_df <- full_df[!is.na(full_df$date_end_accel), ]

difftime_weeks <- difftime(full_df$end_of_fu, full_df$date_end_accel, units = "weeks")

# because the "difftime" function only allows a maximum of units in weeks, we 
# devide by 4.345 to convert to difference months
full_df$fu_time <- as.numeric(difftime_weeks) / 4.345

head(full_df$fu_time) 
#print(full_df$date_start_accel)

hist(full_df$fu_time, main = "survival time in Month")



```


##### Create binary variable for CVD (strokes and MI)

```{r}
# First we create a new column which will indicate if a participant has had: 
## a) a stroke == 1; or not == 0:
stroke_cols <- c('stroke_date', 'ischaemic_stroke_date', 'intracerebral_haemorrhage_date')

# Create the 'stroke' column
full_df$stroke <- ifelse(rowSums(!is.na(full_df[, stroke_cols ])) > 0, 1, 0)


## b) a MI == 1; or not == 0:
MI_cols <- c('myocardial_infarction_date', 'STEMI_date', 'NSTEMI_date')
full_df$MI <- ifelse(rowSums(!is.na(full_df[, MI_cols ])) > 0, 1, 0)


head(full_df$stroke)
head(full_df$MI)

# We now create a composit "CVD" column which indicates if the participants has had a
# stroke and/or MI (==1) or not (==0)
full_df <- full_df %>%
  mutate(CVD = ifelse(stroke == 1 | MI == 1, 1, 0))
```






################ GitHub PA code

# Now link data with acclerometer weekly minutes, for LPA, MVPA, and sedendary behavior.   
- There are 3 diferent accelerometric measures   
  1. PA1 - Vector magnitude based measures as calcuated from time-series of vector magnitude by UKBB 
  2. PA2 - machine-learning derived mesared, for Field ID 1020
  3. PA3 (pending) - Actigraph count-based measure, based on Freedson cutoff 


#Load derived accelerometer from UKBB, PA2 data 
```{r, include=FALSE}
ML_derived_PA2 <- fread("/home/yacine/survival_analysis/PA2_category_1020.csv")
ML_derived_PA2 <- ML_derived_PA2 %>%  dplyr::select(
  eid, 
  "40048-0.0",  
  "40049-0.0",
  "40047-0.0", 
  "40046-0.0", 
  "40044-0.0",  
  "40045-0.0",
  "40043-0.0", 
  "40042-0.0" 
)

# relabel for id, weekly light PA, weekly MVPA, weekly sleep, weekly sedentary, daily light PA, daily MVPA, daily sedentary
names(ML_derived_PA2) <- c("eid", "overall_l", "overall_mv", "overall_sb", "overall_slp", "day_l", "day_mv", "day_sb", "day_slp")

ML_derived_PA2 <- ML_derived_PA2 %>% filter(!is.na(overall_l))

dim(ML_derived_PA2)


ML_derived_PA2$MVPA_min <- ML_derived_PA2$overall_mv * 7 * 24 * 60 



ML_derived_PA2$MVPA_min <- ML_derived_PA2$overall_mv* 7*24*60
ML_derived_PA2$LPA_min <- ML_derived_PA2$overall_l* 7*24*60
ML_derived_PA2$SB_min <- ML_derived_PA2$overall_sb* 7*24*60

colnames(ML_derived_PA2) <- paste(colnames(ML_derived_PA2), "_PA2", sep = "")
colnames(ML_derived_PA2)[1] <- "eid"


```




# Files from parallel processing of time-series (FieldID 90004) to generate ENMO summary of 100,000 ppl, PA1 data 
```{r, include=FALSE, message=FALSE}
enmoTSParallel_PA1 <- list.files(path="/home/yacine/survival_analysis/PA1_parallel_TS_ENMO/outputs", full.names = TRUE) %>% 
  lapply(read_csv) %>% 
  bind_rows 

enmoTSParallel_PA1$MVPA_min <- enmoTSParallel_PA1$MVPA_total* 7*24*60
enmoTSParallel_PA1$LPA_min <- enmoTSParallel_PA1$LPA_total* 7*24*60
enmoTSParallel_PA1$SB_min <- enmoTSParallel_PA1$SB_total* 7*24*60

colnames(enmoTSParallel_PA1) <- paste(colnames(enmoTSParallel_PA1), "_PA1", sep = "")
colnames(enmoTSParallel_PA1)[1] <- "eid"




```



We merge the full_df and PA1 and PA2 df by eids
```{r}
full_df_PA <- inner_join(
  full_df, 
  enmoTSParallel_PA1 %>% 
    dplyr::select(eid, MVPA_min_PA1, LPA_min_PA1, SB_min_PA1),
  by = "eid")

full_df_PA<- inner_join(full_df_PA, 
  ML_derived_PA2 %>% 
    dplyr::select(eid, MVPA_min_PA2, LPA_min_PA2, SB_min_PA2),
  by = "eid")

```




```{r}
sum(full_df_PA$MVPA_min_PA1 > 2500 )
summary(full_df_PA$MVPA_min_PA1)

sum(full_df_PA$MVPA_min_PA2 > 2500 )
summary(full_df_PA$MVPA_min_PA2)


dim(full_df_PA)
full_df_PA <- full_df_PA %>%  filter(MVPA_min_PA2 < 2500)
full_df_PA <- full_df_PA %>%  filter(MVPA_min_PA1 < 2500)
dim(full_df_PA)



full_df_PA$MVPA_Quant_PA1 <- dvmisc::quant_groups(full_df_PA$MVPA_min_PA1, 4) %>% factor()
full_df_PA$MVPA_Quant_PA2 <- dvmisc::quant_groups(full_df_PA$MVPA_min_PA2, 4) %>% factor()





boxplot(MVPA_min_PA1 ~ MVPA_Quant_PA1, data = full_df_PA)
boxplot(MVPA_min_PA1 ~ MVPA_Quant_PA2, data = full_df_PA)


```



```{r}
# Re-order categories 
full_df_PA$cooked_vg %>% table
full_df_PA$cooked_vg <- factor(full_df_PA$cooked_vg)
levels(full_df_PA$cooked_vg)
levels(full_df_PA$cooked_vg) <- list( 'Less than 2 servings a day' ="< 2 servings/day", 'Between 2 and 4 servings a day' = "2 to 4 servings/day", 'More than 4 servings a day' =  "More than 4 servings/day") 
full_df_PA$cooked_vg %>% table


full_df_PA$fresh_fruit %>% table
full_df_PA$fresh_fruit <- factor(full_df_PA$fresh_fruit)
levels(full_df_PA$fresh_fruit)
levels(full_df_PA$fresh_fruit) <- list( 'Less than 2 servings a day' ="< 2 servings/day", 'Between 2 and 4 servings a day' = "2 to 4 servings/day", 'More than 4 servings a day' =  "More than 4 servings/day") 
full_df_PA$fresh_fruit %>% table


<<<<<<< HEAD
full_df_PA$smoking %>% table 
full_df_PA$smoking<- factor(full_df_PA$smoking)
levels(full_df_PA$smoking)
levels(full_df_PA$smoking) <- list( 'Never' = "Never", 
                                    'Previous' = "Previous", 
                                    'Current' =  "Current") 
=======


# Variables to add to the table - this variable cannot be used to select colum,ns now, since cooked_veg is not rendred into summary table 
includeVar <- c(
  "fu_time",
"MVPA_Quant_PA2",
"MVPA_Quant_PA1",
"MVPA_min_PA1",
"stroke",
"MI",
"age_entry_years",
"ethnicity",
"sex",
"education_level",
"tdi_quarters",
"BMI",
"diabetes",
"smoking",
"alcohol",
"processed_meat",
"red_meat",
"oily_fish",
"non_oily_fish",
"fresh_fruit",
"cooked_vg"
  )


d <-  full_df_incl_PA1_PA2 %>% dplyr::select(all_of(includeVar))
d <- d %>%  
  dplyr::mutate(MVPA_Quant_PA1 = factor(MVPA_Quant_PA1),
                MVPA_Quant_PA2 = factor(MVPA_Quant_PA2))

>>>>>>> 8a479adab37ae4000792ef2d9d84a445a98acc5e


# Obesity category names can be shortened 
# Note that all vars , except alcohol, do not know and unkjnan shoudl be meet with exclusion 
dt <- full_df_PA %>%  dplyr::select(
  fu_time, 
  MVPA_Quant_PA2,  
  MVPA_Quant_PA1,  
  MVPA_min_PA1, 
  MVPA_min_PA2, 
  MI, 
  stroke,
  age_entry_years, 
  ethnicity, 
  sex, 
  education_level,  
  tdi_quarters,
  BMI,
  diabetes, 
  smoking, 
  alcohol, 
  processed_meat, 
  red_meat,
  oily_fish,
  non_oily_fish,
  fresh_fruit,
  cooked_vg
  )



# Convert varaible names into deescriptive tabl enames 
varNameDf <- data.frame(
<<<<<<< HEAD
  varName = c("Myocardial Infarction",
              "Age",
              "MVPA min/week",
              "Race", 
              "Type II Diabetes", 
              "Deprivation index", 
              "Education", 
              "Smoking", 
              "Alcohol", 
              "Processed meat", 
              "Fresh fruit", 
              "Cooked vegetables", 
              "Oily fish", 
              "Sex", 
              "Follow_up time"
  ), 
  ChartName = c("MI", 
              "age_entry_years",
              "MVPA_min_PA1", 
              "ethnicity", 
              "diabetes", 
              "tdi_quarters", 
              "education_level", 
              "smoking", 
              "alcohol", 
              "processed_meat", 
              "fresh_fruit", 
              "cooked_vg", 
              "oily_fish", 
              "sex", 
              "fu_time")
=======
  varName = c("MVPA_MIN_PA1",
    "MI",
    "STROKE",
    "AGE_ENTRY_YEARS",
    "ETHNICITY",
    "SEX",
    "EDUCATION_LEVEL",
    "TDI_QUARTERS",
    "BMI",
    "DIABETES",
    "SMOKING",
    "ALCOHOL",
    "PROCESSED_MEAT",
    "RED_MEAT",
    "OILY_FISH",
    "NON_OILY_FISH",
    "FRESH_FRUIT",
    "COOKED_VG"
  ), 
  ChartName = c("MVPA_min_PA1",
                "MI",
                "stroke",
                "age_entry_years",
                "ethnicity",
                "sex",
                "education_level",
                "tdi_quarters",
                "BMI",
                "diabetes",
                "smoking",
                "alcohol",
                "processed_meat",
                "red_meat",
                "oily_fish",
                "non_oily_fish",
                "fresh_fruit",
                "cooked_vg")
>>>>>>> 8a479adab37ae4000792ef2d9d84a445a98acc5e
)


var_names <- varNameDf %>% 
    select(varName, ChartName) %>% 
    deframe()

dt <- dt %>% 
    rename(!!!var_names)


tb <- dt %>%
  tbl_summary(
    by = MVPA_Quant_PA2,  
    #statistic = all_continuous() ~ "{mean} ({sd})"
    #digits = all_continuous() ~ 1,
<<<<<<< HEAD
=======
  label = c(
     varNameDf[1, "ChartName"] ~ varNameDf[1, "varName"],
        varNameDf[2, "ChartName"] ~ varNameDf[2, "varName"],
        varNameDf[3, "ChartName"] ~ varNameDf[3, "varName"],
        varNameDf[4, "ChartName"] ~ varNameDf[4, "varName"],
        varNameDf[5, "ChartName"] ~ varNameDf[5, "varName"],
        varNameDf[6, "ChartName"] ~ varNameDf[6, "varName"],
        varNameDf[7, "ChartName"] ~ varNameDf[7, "varName"],
        varNameDf[8, "ChartName"] ~ varNameDf[8, "varName"],
        varNameDf[9, "ChartName"] ~ varNameDf[9, "varName"],
        varNameDf[10, "ChartName"] ~ varNameDf[10, "varName"],
        varNameDf[11, "ChartName"] ~ varNameDf[11, "varName"],
        varNameDf[12, "ChartName"] ~ varNameDf[12, "varName"],
        varNameDf[13, "ChartName"] ~ varNameDf[13, "varName"],
        varNameDf[14, "ChartName"] ~ varNameDf[14, "varName"],
        varNameDf[15, "ChartName"] ~ varNameDf[15, "varName"],
        varNameDf[16, "ChartName"] ~ varNameDf[16, "varName"],
        varNameDf[17, "ChartName"] ~ varNameDf[17, "varName"],
        varNameDf[18, "ChartName"] ~ varNameDf[18, "varName"]
)
>>>>>>> 8a479adab37ae4000792ef2d9d84a445a98acc5e
  ) %>% 
  as_flex_table()

# cOMPRESSED table 
theme_gtsummary_compact() 

# Landscape (word format)
sect_properties <- prop_section(
  page_size = page_size(
  orient = "landscape",
  #width = 8, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar(0) 
  )


print(tb)

# Save 
flextable::save_as_docx("Distribution of participant characteristics by Qurtile of Moderate_Vigorous Physical Activity" = tb,  
               path = "testTable.docx", 
               pr_section = sect_properties
               )  
```


```{r}
#pairs(inl)

library(ggcorrplot)
model.matrix(~0+., data=dt) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag=FALSE, type="lower", lab=F, lab_size=2)




```





##############################################################



##########
Kaplan_Meier:
  - 2 groups: PA1 and PA2 or Stroke and MI?
  - d_stroke == 1; d_MI == 1
  - end_of_fu = censoring time
  
  
  
  
## Associations with risk of incident cardiovascular disease 

In the data preparation step, we added an event status indicator at exit and a follow-up time variable. Using these, we can run a Cox model to associate overall activity with risk of incident cardiovascular disease. We'll start by using time-on-study as the timescale and set it up using the 'survival' package in R. We'll also adjust for various possible confounding variables (following the confounders used by [Ramakrishnan et al.](https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.1003487)):


### Regression 
```{r}

<<<<<<< HEAD
fitDf <- coxph(Surv(`Follow_up time`,`Myocardial Infarction`) ~ 
               Sex + 
              `Age` +
              `MVPA min/week`+
              `Race` + 
              `Type II Diabetes`+ 
              `Deprivation index`+ 
              `Education`+ 
              `Smoking`+ 
              `Alcohol`+ 
              `Processed meat`+ 
              `Fresh fruit`+ 
              `Cooked vegetables`+ 
              `Oily fish`, 
data = dt)
=======
#km_stroke <- coxph(
  #Surv(fu_time, stroke) ~ overall_activity_quarters + age_entry_years + sex + ethnicity + tdi_quarters + education_level + smoking + alcohol,
  #data = full_df_incl_PA1_PA2)
#install.packages("ggsurvplot")
#library(ggsurvplot)

varNameDf <- data.frame(
  varName = c("MVPA_Quant_PA2", 
              "AGE_ENTRY_YEARS", 
              "ETHNICITY", 
              "SEX", 
              "EDUCATION_LEVEL", 
              "TDI_QUARTERS", 
              "BMI", 
              "DIABETES", 
              "SMOKING", 
              "ALCOHOL", 
              "PROCESSED_MEAT", 
              "RED_MEAT", 
              "OILY_FISH", 
              "NON_OILY_FISH", 
              "FRESH_FRUIT", 
              "COOKED_VG"), 
  ChartName = c("MVPA_Quant_PA2",
  "age_entry_years",
  "ethnicity",
  "sex",
  "education_level",
  "tdi_quarters",
  "BMI",
  "diabetes",
  "smoking",
  "alcohol",
  "processed_meat",
  "red_meat",
  "oily_fish",
  "non_oily_fish",
  "fresh_fruit",
  "cooked_vg")
)


fitDf <- coxph(Surv(fu_time, MI) ~ 
                MVPA_Quant_PA2 +
                age_entry_years +
                ethnicity +
                sex +
                education_level +
                tdi_quarters +
                BMI +
                diabetes +
                smoking +
                alcohol +
                processed_meat +
                red_meat +
                oily_fish +
                non_oily_fish +
                fresh_fruit +
                cooked_vg,
                data = full_df_incl_PA1_PA2)
#ggsurvplot(survfit(res.cox), color = "#2E9FDF",
           #ggtheme = theme_minimal())
>>>>>>> 8a479adab37ae4000792ef2d9d84a445a98acc5e

summary(fitDf)
tbl_regression(fitDf, exponentiate = TRUE)



<<<<<<< HEAD
=======
tbl_regression(fitDf, exponentiate = TRUE, 
              label = c(
        varNameDf[1, "ChartName"] ~ varNameDf[1, "varName"],
        varNameDf[2, "ChartName"] ~ varNameDf[2, "varName"],
        varNameDf[3, "ChartName"] ~ varNameDf[3, "varName"],
        varNameDf[4, "ChartName"] ~ varNameDf[4, "varName"],
        varNameDf[5, "ChartName"] ~ varNameDf[5, "varName"],
        varNameDf[6, "ChartName"] ~ varNameDf[6, "varName"],
        varNameDf[7, "ChartName"] ~ varNameDf[7, "varName"],
        varNameDf[8, "ChartName"] ~ varNameDf[8, "varName"],
        varNameDf[9, "ChartName"] ~ varNameDf[9, "varName"],
        varNameDf[10, "ChartName"] ~ varNameDf[10, "varName"],
        varNameDf[11, "ChartName"] ~ varNameDf[11, "varName"],
        varNameDf[12, "ChartName"] ~ varNameDf[12, "varName"],
        varNameDf[13, "ChartName"] ~ varNameDf[13, "varName"],
        varNameDf[14, "ChartName"] ~ varNameDf[14, "varName"],
        varNameDf[15, "ChartName"] ~ varNameDf[15, "varName"],
        varNameDf[16, "ChartName"] ~ varNameDf[16, "varName"]
   )
)


>>>>>>> 8a479adab37ae4000792ef2d9d84a445a98acc5e


```
  




