---
title: "UKBB Data Exclusion"
author: "Yalap"
output: rmarkdown::github_document
---

We load and/or install the packages we will be using:

```{r, include=FALSE}
rm(list=ls())
library(dplyr)
library(readr)
library(data.table)
#install.packages(tidyverse)
library(tidyverse)
```

We import all data for participants with accelerometer data:

```{r}
# Change the locations to your user folder data files
#full_df <- read.csv("/home/hiroshi/projects/UKBB/Yacine_extract/final_df_full.csv")
full_df <- read.csv("/home/yacine/UKBB_beluga/final_df_full.csv") # copy in /home/yacine/survival_analysis

```

############ Data prep intro ############

We convert "-1" and "-3" values to NAs for the self-reported PA values for 
instance 0
```{r}




full_df <- full_df %>%
  mutate(
    duration_moderate_activity_0 = na_if(as.numeric(as.character(duration_moderate_activity_0)), -1),
    duration_moderate_activity_0 = na_if(duration_moderate_activity_0, -3),
    
    duration_vigorous_activity_0 = na_if(as.numeric(as.character(duration_vigorous_activity_0)), -1),
    duration_vigorous_activity_0 = na_if(duration_vigorous_activity_0, -3),
    
    #duration_heavy_DIY_0 = na_if(as.numeric(as.character(duration_heavy_DIY_0)), -1),
    #duration_heavy_DIY_0 = na_if(duration_heavy_DIY_0, -3),
    
   # duration_light_DIY_0 = na_if(as.numeric(as.character(duration_light_DIY_0)), -1),
   # duration_light_DIY_0 = na_if(duration_light_DIY_0, -3),
  )

```


We convert "-1" and "-3" values to NAs for the self-reported PA values for 
instances 1 (2012-2013) and instances 2 (2014+)
```{r}
#full_df <- full_df %>%
  #mutate(
    #duration_moderate_activity_1 = na_if(as.numeric(as.character(duration_moderate_activity_1)), -1),
    #duration_moderate_activity_1 = na_if(duration_moderate_activity_1, -3),
    #duration_moderate_activity_2 = na_if(as.numeric(as.character(duration_moderate_activity_2)), -1),
    #duration_moderate_activity_2 = na_if(duration_moderate_activity_2, -3),    
    
    #duration_vigorous_activity_1 = na_if(as.numeric(as.character(duration_vigorous_activity_1)), -1),
    #duration_vigorous_activity_1 = na_if(duration_vigorous_activity_1, -3),
    #duration_vigorous_activity_2 = na_if(as.numeric(as.character(duration_vigorous_activity_2)), -1),
    #duration_vigorous_activity_2 = na_if(duration_vigorous_activity_2, -3),
    
    
    #duration_heavy_DIY_1 = na_if(as.numeric(as.character(duration_heavy_DIY_1)), -1),
    #duration_heavy_DIY_1 = na_if(duration_heavy_DIY_1, -3),
    #duration_heavy_DIY_2 = na_if(as.numeric(as.character(duration_heavy_DIY_2)), -1),
    #duration_heavy_DIY_2 = na_if(duration_heavy_DIY_2, -3),
    
    #duration_light_DIY_1 = na_if(as.numeric(as.character(duration_light_DIY_1)), -1),
    #duration_light_DIY_1 = na_if(duration_light_DIY_1, -3),
    #duration_light_DIY_2 = na_if(as.numeric(as.character(duration_light_DIY_2)), -1),
    #duration_light_DIY_2 = na_if(duration_light_DIY_2, -3)    
  #)
```

We combine the self reported PA data for instances 0
```{r}

#full_df <- full_df %>%
 #mutate(
    #duration_moderate_activity = coalesce(duration_moderate_activity_1, duration_moderate_activity_2),
    #duration_vigorous_activity = coalesce(duration_vigorous_activity_1, duration_vigorous_activity_2),
    #duration_heavy_DIY = coalesce(duration_heavy_DIY_1, duration_heavy_DIY_2),
    #duration_light_DIY = coalesce(duration_light_DIY_1, duration_light_DIY_2)
  #)
```

### PA4 prep:
Combining the duration_moderate_activity_0 and duration_vigorous_activity_0
we create the "MVPA_min_PA4" by doing the sum of both columns

```{r}

full_df <- full_df %>%
  mutate(
    MVPA_min_PA4 = coalesce(duration_vigorous_activity_0) + coalesce(duration_moderate_activity_0)
  )

```


Combine green/blue spaces data into one columns
```{r}

full_df <- full_df %>%
  mutate(water_300m = coalesce(water_300m_0, water_300m_1))

full_df <- full_df %>%
  mutate(greenspace_300m = coalesce(greenspace_300m_0, greenspace_300m_1))

full_df <- full_df %>%
  mutate(naturalenv_300m = coalesce(naturalenv_300m_0, naturalenv_300m_1))

### I did not include garden

```

######## REMOVE UNNECESSARY COLUMNS
After looking at the data, I realized that some columns that we extracted (for
multiple instances) provided more "NA" data than actual data, so I removed these 
columns:
```{r}

full_df <- full_df %>%
  dplyr::select(-diabetes_1, -diabetes_2, -diabetes_3, 
         -processed_meat_1, -processed_meat_2, -processed_meat_3, -X191.0.0.1, -duration_moderate_activity_1, -duration_moderate_activity_2, -duration_moderate_activity_3, -duration_vigorous_activity_1, -duration_vigorous_activity_2, -duration_vigorous_activity_3, -duration_heavy_DIY_1, -duration_heavy_DIY_2, -duration_heavy_DIY_3, -duration_light_DIY_1, -duration_light_DIY_2, -duration_light_DIY_3, -domgarden_1000m_0, -domgarden_300m_0, -greenspace_1000m_0, -domgarden_1000m_1, -domgarden_300m_1, -greenspace_1000m_1, -water_1000m_0, -water_1000m_1, -dist_coast_0, dist_coast_1)

```


########## Exclusions: these following commands will only keep the data we are ##########  
            not seeking to exclude

First, we would like to keep count of the number of participants that are excluded
at each step. So we create this function that will allow us to do so:
```{r}
count_exclusions <- function(df_before, df_after) {
  n_before <- nrow(df_before)
  n_after <- nrow(df_after)
  excluded <- n_before - n_after
  print(paste("Participants excluded:", excluded))
  return(list(df = df_after, excluded = excluded))
}

initial_n <- nrow(full_df) # number of participants at the beginning
print(initial_n)
```

## Exclusion 1: Low quality accelerometer data

```{r}
# we want to count how many people were excluded in this section
total_excluded_lq <- 0

# Added by Hiroshi
df_before <- full_df
full_df <- full_df[full_df$accel_raw != "", ]

result <- count_exclusions(df_before, full_df)
total_excluded_lq <- total_excluded_lq + result$excluded
```

### 1.1. Devices poorly calibrated
```{r}
df_before <- full_df
full_df <- full_df[full_df$quality_good_calibration == "Yes", ]
result <- count_exclusions(df_before, full_df)
total_excluded_lq <- total_excluded_lq + result$excluded
```

### 1.2. Participants for whom \>1% of values were clipped (fell outside the 
sensor's range) before or after calibration:
```{r}
df_before <- full_df
full_df <- full_df[(full_df$clips_before_cal < 0.01 * full_df$total_reads) & (full_df$clips_after_cal < 0.01 * full_df$total_reads), ]
result <- count_exclusions(df_before, full_df)
total_excluded_lq <- total_excluded_lq + result$excluded
```

### 1.3 Insufficient wear time
```{r}
df_before <- full_df
full_df <- full_df[full_df$quality_good_wear_time == "Yes", ]# OxWearable note: Note that this has already been calculated in UKB, we don't need to manually calculate it: https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=90015
result <- count_exclusions(df_before, full_df)
total_excluded_lq <- total_excluded_lq + result$excluded
```

### 1.4 Unrealistically high acceleration values
```{r}
df_before <- full_df
full_df <- full_df[full_df$overall_activity < 100, ]
result <- count_exclusions(df_before, full_df)
total_excluded_lq <- total_excluded_lq + result$excluded
```

```{r}
print(paste("Total participants excluded due to not being in acceleromete subcohort or low accelerometer data quality:", total_excluded_lq))
```


## Exclusion 2: Reported CVD

```{r}
# we extract the CVD date columns
date_columns <- c(
  'stroke_date', 'ischaemic_stroke_date', 'intracerebral_haemorrhage_date', 
  'myocardial_infarction_date', 'STEMI_date', 'NSTEMI_date' , 'date_end_accel', 'date_start_accel' 
)

# we format the date columns
for (col in date_columns) {
  full_df[[col]] <- as.Date(full_df[[col]], format="%Y-%m-%d")
}
```


### Removal of CVD cases based on the baseline date of acceleroemter for each person:

```{r}
# Max date of diagnosis (hospital record update) is 2022 Dec. 
summary(full_df$stroke_date) #originally 1140 excluded
summary(full_df$ischaemic_stroke_date)
summary(full_df$intracerebral_haemorrhage_date)
summary(full_df$myocardial_infarction_date) # orignally 1796  excluded
summary(full_df$STEMI_date)
summary(full_df$NSTEMI_date)
df_beforeExclusion <- full_df %>% dplyr::select(all_of(c(date_columns, "eid")))

# cut-off date = 2013-01-01 because we do not want to keep participants that reported
# having a CVD < 2013
cut_off_date <- "2013-01-01"
cut_off_date_personal <- full_df$date_end_accel 
full_df$indicator_drop_stroke <- 1
full_df$indicator_drop_mi <- 1
full_df[is.na(full_df[["stroke_date"]]) | full_df[["stroke_date"]] >= as.Date(cut_off_date_personal), "indicator_drop_stroke"] <- 0 
full_df[is.na(full_df[["myocardial_infarction_date"]]) | full_df[["myocardial_infarction_date"]] >= as.Date(cut_off_date_personal), "indicator_drop_mi"] <- 0

sum(full_df$indicator_drop_stroke) # to remove. 1251
sum(full_df$indicator_drop_mi)  # to remove, 2181
full_df %>% dplyr::filter(indicator_drop_mi == 1 & indicator_drop_stroke==1) %>% nrow 

# 
print("Number of people excluded based on the exact date of acclerometer is \n")
full_df %>% filter(indicator_drop_mi == 1 | indicator_drop_stroke==1) %>% dim()

# Create data frame after exclusion 
full_df_CVD <- full_df %>% # 3315 removed by this definition 
  filter(indicator_drop_mi == 0 & indicator_drop_stroke==0)

# Check dates 
min(full_df_CVD$stroke_date, na.rm = T)
max(full_df_CVD$stroke_date, na.rm = T)
min(full_df_CVD$myocardial_infarction_date, na.rm = T)
max(full_df_CVD$myocardial_infarction_date, na.rm = T)

# Confirm the last date of diagnosis among those who are removed
df_beforeExclusion %>% 
  anti_join(full_df_CVD, by =join_by(eid)) %>% 
  select(all_of(date_columns)) %>% 
  apply(1, FUN = min, na.rm=T) %>% max()

# Confirm the # of excluded # 2936
df_beforeExclusion %>% 
    anti_join(full_df_CVD, by =join_by(eid)) %>% 
    select(all_of(date_columns)) %>% 
    apply(1, FUN = min, na.rm=T) %>% length
#format(as.Date(.), "%Y-%m") %>% unique()  %>%  sort

```


### Find the last date of hospital records and make exclusion based on the dates:

```{r}
# we create an empty list that will be filled with the number of excluded participants
exclusions <- list()

# we create a function which will allow us to count the number of excluded participants
count_and_filter_dates <- function(df, date_column, cut_off_date) {
  initial_count <- nrow(df)
  df <- df[is.na(df[[date_column]]) | df[[date_column]] >= as.Date(cut_off_date), ]
  final_count <- nrow(df)
  excluded_count <- initial_count - final_count # number of excluded rows (participants)
  print(paste("Excluded", date_column, excluded_count))
  return(list(df = df, excluded_count = excluded_count))
}

# create loop to count number of participants excluded for each CVD
total_excluded_cvd <- 0
for (col in date_columns) {
  result <- count_and_filter_dates(full_df, col, cut_off_date)
  full_df <- result$df
  exclusions[[paste("Excluded based on", col)]] <- result$excluded_count
  total_excluded_cvd <- total_excluded_cvd + result$excluded_count
}

# Print total exclusions for step 2
print(paste("Total participants excluded because of CVD:", total_excluded_cvd)) #originally 2936

```


## Exclusion 3: Remove participants for whom we have lost follow-up data

```{r}
df_before <- full_df
full_df <- full_df[full_df$date_lost_followup == "", ]
result <- count_exclusions(df_before, full_df)
total_excluded_lfu <- result$excluded

# Print total exclusions for step 3
print(paste("Total participants excluded in step 3:", total_excluded_lfu))

```

## Final count after exclusion steps 1 to 3:
```{r}
final_n <- nrow(full_df)
print(paste("After applying exclusions 1 to 3, we end up with a sample of n =", final_n," participants"))
```


## Save df with all excluded participants:

```{r}
#write.csv(full_df, "/home/yacine/UKBB_beluga/df_exc.csv", row.names = FALSE)
```



########## Data Recoding ########## 

We recode certain variables: 
### Note: all the data coding/mapping was an extra step we had to do ###

1. Ethnicity:

```{r}

##first we create a dictionnary mapping for the ethnicity codes
## *** important: this is not a step that was shown by OxWearables. We had to this
## ourselves. The codes are provided here: https://biobank.ctsu.ox.ac.uk/crystal/coding.cgi?id=1001

ethnicity_coding <- c(
  "1" = "White",
  "1001" = "British",
  "2001" = "White and Black Caribbean",
  "3001" = "Indian",
  "4001" = "Caribbean",
  "2" = "Mixed",
  "1002" = "Irish",
  "2002" = "White and Black African",
  "3002" = "Pakistani",
  "4002" = "African",
  "3" = "Asian or Asian British",
  "1003" = "Any other white background",
  "2003" = "White and Asian",
  "3003" = "Bangladeshi",
  "4003" = "Any other Black background",
  "4" = "Black or Black British",
  "2004" = "Any other mixed background",
  "3004" = "Any other Asian background",
  "5" = "Chinese",
  "6" = "Other ethnic group",
  "-1" = "Do not know",
  "-3" = "Prefer not to answer"
)

## we replace the codes by their meaning:
full_df <- full_df %>%
  mutate(ethnicity_raw = recode(as.character(ethnicity_raw), !!!ethnicity_coding))


## 2. now we can proceed with the recoding suggested by OxWearables:
full_df$ethnicity <-
  plyr::revalue(
    full_df$ethnicity_raw,
    c(
      "British" = "White",
      "Any other white background" = "White",
      "Irish" = "White",
      "White and Asian" = "Nonwhite",
      "Caribbean" = "Nonwhite",
      "Chinese"   = "Nonwhite",
      "Pakistani"  = "Nonwhite",
      "White and Black African" = "Nonwhite",
      "Other ethnic group"  = "Nonwhite",
      "Any other mixed background" = "Nonwhite",
      "African"    = "Nonwhite",
      "White and Black Caribbean" = "Nonwhite",
      "Prefer not to answer" = NA,
      "Indian"  = "Nonwhite",
      "White" = "White",
      "Do not know" = NA,
      "Any other Black background" = "Nonwhite",
      "Any other Asian background"  = "Nonwhite",
      "Bangladeshi"  = "Nonwhite",
      "Mixed"  = "Nonwhite",
      "Asian or Asian British"  = "Nonwhite",
      "Black or Black British"  = "Nonwhite"
    )
  )



table(full_df$ethnicity_raw, full_df$ethnicity, useNA = "always")
```


2. Smoking and tdi:

```{r}
# Townsend Index:
full_df$tdi <- full_df$tdi_raw

# Smoking
## 1. Data coding for smoking: https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=90
smoking_coding <- c(
  "0" = "Never",
  "1" =	"Previous",
  "2" =	"Current",
  "-3" = "Prefer not to answer"
)

## we replace the codes by their meaning:
full_df <- full_df %>%
  mutate(smoking = recode(as.character(smoking_raw), !!!smoking_coding))

## 2. OxWearable recoding:
full_df$smoking <-
  plyr::revalue(full_df$smoking, replace = c("Prefer not to answer" = NA)) 

table(full_df$smoking, full_df$smoking_raw, useNA = "always")
```



3. Fish and meats:

```{r}
##### Recoding we added

# Recording for oily and non-oily fish and processed meats
meat_coding <- c(
  "1" = "Less than once a week",
  "2" =	"Once a week",
  "3" =	"2-4 times a week",
  "4" =	"5-6 times a week",
  "5" =	"Once or more daily",
  "0" =	"Never",
  "-1" = "Do not know",
  "-3" = "Prefer not to answer"
)

## we replace the codes by their meaning:
### oily fish:
full_df$oily_fish_raw <- full_df$oily_fish  
full_df <- full_df %>%
  mutate(oily_fish = recode(as.character(oily_fish), !!!meat_coding))

### non-oily fish:
full_df$non_oily_fish_raw <- full_df$non_oily_fish  
full_df <- full_df %>%
  mutate(non_oily_fish = recode(as.character(non_oily_fish  ), !!!meat_coding))

### processed meats:
full_df <- full_df %>%
  mutate(processed_meat_0 = recode(as.character(processed_meat_0), !!!meat_coding ))

table(full_df$oily_fish, full_df$oily_fish_raw, useNA = "always" ) # Yacine comment: remove? There twice



# Oily fish
full_df <- full_df %>%
  mutate(oily_fish = case_when(
    oily_fish == "2-4 times a week" ~ "2-4 times a week",
    oily_fish == "Prefer not to answer" ~ "Prefer not to answer",
    oily_fish == "Do not know" ~ "Do not know",
    oily_fish %in% c("Once a week", "Less than once a week", "Never") ~ "Less than 2 times a week",
    oily_fish %in% c("5-6 times a week", "Once or more daily") ~ "More than 4 times a week",
    TRUE ~ NA_character_
  ))

table(full_df$oily_fish, full_df$oily_fish_raw, useNA = "always" )

# Non-oily fish
full_df <- full_df %>%
  mutate(non_oily_fish = case_when(
    non_oily_fish == "2-4 times a week" ~ "2-4 times a week",
    non_oily_fish == "Prefer not to answer" ~ "Prefer not to answer",
    non_oily_fish == "Do not know" ~ "Do not know",
    non_oily_fish %in% c("Once a week", "Less than once a week", "Never") ~ "Less than 2 times a week",
    non_oily_fish %in% c("5-6 times a week", "Once or more daily") ~ "More than 4 times a week",
    TRUE ~ NA_character_
  ))

table(full_df$non_oily_fish, full_df$non_oily_fish_raw, useNA = "always" )


# Processed meats
full_df$processed_meat_raw <- full_df$processed_meat_0
full_df <- full_df %>%
  mutate(processed_meat_0 = case_when(
    processed_meat_0 == "2-4 times a week" ~ "2-4 times a week",
    processed_meat_0 == "Prefer not to answer" ~ "Prefer not to answer",
    processed_meat_0 == "Do not know" ~ "Do not know",
    processed_meat_0 %in% c("Once a week", "Less than once a week", "Never") ~ "Less than 2 times a week",
    processed_meat_0 %in% c("5-6 times a week", "Once or more daily") ~ "More than 4 times a week",
    TRUE ~ NA_character_
  ))

# rename processed_meat variable
full_df$processed_meat <- full_df$processed_meat_0
full_df <- full_df %>%
  select(-processed_meat_0)

table(full_df$processed_meat, full_df$processed_meat_raw, useNA = "always" )
```

4. Red meats:
* we need to create the red_meat column as the UKBiobank data only provides data 
for different types of red meat separately (i.e. lamb, beef and pork) *

```{r}

## we replace "Do not know"s and "Prefer not to say"s by NAs - we will drop
## them later on anyways:
full_df <- full_df %>%
  mutate(across(c(beef, lamb, pork), ~ na_if(., -1))) %>%
  mutate(across(c(beef, lamb, pork), ~ na_if(., -3)))

## we create the red meat column by adding all the values for beef, lamb and pork: 
full_df$red_meat <- rowSums(full_df[c("beef", "lamb", "pork")], na.rm = TRUE)

table(full_df$red_meat) # visualize all the value that take red_meat

## we create coding for red meats:
red_meat_coding <- c(
  "0" = "Never",
  "1"  = "Less than once a week",
  "2"  = "Once a week",
  "3"  = "2-4 times a week",
  "4"  = "5-6 times a week",
  "5"  = "Once or more daily",
  "6"  = "Once or more daily",
  "7"  = "Once or more daily",
  "8"  = "Once or more daily",
  "9"  = "Once or more daily",
  "10" = "Once or more daily",
  "11" = "Once or more daily",
  "12" = "Once or more daily",
  "13" = "Once or more daily",
  "15" = "Once or more daily"
)

## we replace the coding by their meaning:
full_df$red_meat_raw <- full_df$red_meat
full_df <- full_df %>%
  mutate(red_meat = recode(as.character(red_meat), !!!red_meat_coding))

# Red meat
full_df <- full_df %>%
  mutate(red_meat = case_when(
    red_meat == "2-4 times a week" ~ "2-4 times a week",
    red_meat %in% c("Once a week", "Less than once a week", "Never") ~ "Less than 2 times a week",
    red_meat %in% c("5-6 times a week", "Once or more daily") ~ "More than 4 times a week",
    TRUE ~ NA_character_
  ))

table(full_df$red_meat, full_df$red_meat_raw, useNA = "always" )


```



5. Education:

```{r}
#### Note: OxWearable use "age education"; we are using "qualifications education"
## so this is our recoding
education_coding <- c(
  "1" = "College or University degree",
  "2" =	"A levels/AS levels or equivalent",
  "3" =	"O levels/GCSEs or equivalent",
  "4" =	"CSEs or equivalent",
  "5" =	"NVQ or HND or HNC or equivalent",
  "6" =	"Other professional qualifications eg: nursing, teaching",
  "-7" = "None of the above",
  "-3" = "Prefer not to answer"
)

## we replace the education codes by their meaning:
full_df <- full_df %>%
  mutate(education_level = recode(as.character(education_raw), !!!education_coding))

## we recode again to have less education classifications
full_df <- full_df %>%
  mutate(education_level = case_when(
    education_level == "College or University degree" ~ "College or University degree",
    education_level %in% c("A levels/AS levels or equivalent", 
                         "NVQ or HND or HNC or equivalent", 
                         "Other professional qualifications eg: nursing, teaching") ~ 
                         "A levels/AS levels or equivalent, NVQ or HND or HNC or equivalent, other professional qualifications",
    education_level %in% c("O levels/GCSEs or equivalent", "CSEs or equivalent") ~ 
                         "O levels/GCSEs or equivalent, CSEs or equivalent",
    education_level == "None of the above" ~ "None of the above",
    education_level == "Prefer not to answer" ~ "Prefer not to answer",
    TRUE ~ NA_character_
  ))


table(full_df$education_level, full_df$education_raw, useNA = "always")
```



6. BMI:

```{r}
full_df$BMI <- full_df$BMI_raw

## recode the BMI variable 
full_df <- full_df %>%
  mutate(BMI = case_when(
    BMI < 18.5 ~ "Underweight (< 18.5 kg/m2)",
    BMI >= 18.5 & BMI < 25 ~ "Normal (18.5 kg/m2 to < 25 kg/m2)",
    BMI >= 25 & BMI < 30 ~ "Overweight (25 kg/m2 to < 30 kg/m2)",
    BMI >= 30 ~ "Obesity Class I, II or III (> 30 kg/m2)",
    TRUE ~ NA_character_  
  ))

boxplot(BMI_raw ~ BMI, data = full_df)
```


7. Fruit and veggies:

```{r}
# Hiroshi Comment: Zero serving is treated as missing - it should not be. ALso, here you have a varaible in mix of discrete number (integer) and character, and youa re applying oepration for integer. Is this char, integer, factor variable? 
# Hiroshi Comment Zero serving is treated as missing - it should not be
## recode the fruit variables to have less classifications
full_df$fresh_fruit_raw <- full_df$fresh_fruit

full_df <- full_df %>%
  mutate(fresh_fruit = case_when(
    fresh_fruit %in% c(-10, 0) | (fresh_fruit > 0 & fresh_fruit < 2) ~ "< 2 servings/day",
    fresh_fruit == -1 ~ "Do not know",
    fresh_fruit == -3 ~ "Prefer not to answer",
    fresh_fruit >= 2 & fresh_fruit <= 4 ~ "2 to 4 servings/day",
    fresh_fruit > 4 ~ "More than 4 servings/day"
  ))
 

# Veggies:
full_df$cooked_vg_raw <- full_df$cooked_vg

full_df <- full_df %>%
  mutate(cooked_vg = case_when(
    cooked_vg %in% c(-10, 0) | (cooked_vg > 0 & cooked_vg < 2) ~ "< 2 servings/day",
    cooked_vg == -1 ~ "Do not know",
    cooked_vg == -3 ~ "Prefer not to answer",
    cooked_vg >= 2 & cooked_vg <= 4 ~ "2 to 4 servings/day",
    cooked_vg > 4 ~ "More than 4 servings/day"
  ))



table(full_df$fresh_fruit, full_df$fresh_fruit_raw, useNA = "always")
table(full_df$cooked_vg, full_df$cooked_vg_raw, useNA = "always")



```


8. Alcohol:

```{r}
# Alcohol
## 1. Data coding for alochol: https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=100402
alcohol_coding <- c(
  "1" = "Daily or almost daily",
  "2" =	"Three or four times a week",
  "3" =	"Once or twice a week",
  "4" =	"One to three times a month",
  "5" =	"Special occasions only",
  "6" =	"Never",
  "-3" = "Prefer not to answer"
)

## we replace the codes by their meaning:
full_df <- full_df %>%
  mutate(alcohol = recode(as.character(alcohol_raw), !!!alcohol_coding))

## we recode alcohol to have less classifications
full_df <- full_df %>%
  mutate(alcohol = case_when(
    alcohol == "Daily or almost daily" ~ "Daily or almost daily",
    alcohol == "Never" ~ "Never",
    alcohol == "Once or twice a week" ~ "Once or twice a week",
    alcohol == "Three or four times a week" ~ "Three or four times a week",
    alcohol %in% c("One to three times a month", "Special occasions only") ~ "Less than once a week",
    alcohol == "Prefer not to answer" ~ "Prefer not to answer",
    TRUE ~ NA_character_
  ))


table(full_df$alcohol, full_df$alcohol_raw, useNA = "always")
```


9. Diabetes and other comorbidities:

```{r}
# Diabetes
diabetes_coding <- c(
  "0" = "Yes",
  "1" = "No",
  "-1" = "Do not know",
  "-3" = "Prefer not to answer"
)

full_df$diabetes_raw <- full_df$diabetes_0

## we replace the codes by their meaning:
full_df <- full_df %>%
  mutate(diabetes = recode(as.character(diabetes_0), !!!diabetes_coding))

full_df <- full_df %>%
  select(-diabetes_0)

table(full_df$diabetes_raw, full_df$diabetes, useNA = "always")


```


10. Sex:

```{r}
# recode binary variables 
full_df <- full_df %>%
  mutate(sex = recode(sex, `1` = "Male", `0` = "Female"))

table(full_df$sex)
```

11. Average Houselhold Gross Income:

```{r}
income_coding <- c(
  "1" =	"Less than 18,000",
  "2" =	"18,000 to 30,999",
  "3" =	"31,000 to 51,999",
  "4" =	"52,000 to 100,000",
  "5" =	"Greater than 100,000",
  "-1" = "Do not know/Prefer not to answer",
  "-3" = "Do not know/Prefer not to answer"
)

full_df$avg_hh_income_raw <- full_df$avg_hh_income # to keep an exemple of coding

## we replace the codes by their meaning:
full_df <- full_df %>%
  mutate(avg_hh_income = recode(as.character(avg_hh_income_raw), !!!income_coding))


# count how many participants are in each average income ranges
count_avg_hh_income <- full_df %>%
  count(avg_hh_income)

print(count_avg_hh_income)

```





############ Final exclusion ############

## Exclusion 4: Missing covariate data

### 4.1 We remove rows for participants that have missing data for ethnicity, education, sex, birth date, smoking status, alcohol consumption, Townsend Deprivation Index; oily fish, fresh fruit, cooked veggies or processed meat consumption, and BMI.

```{r}

df_before <- full_df

keep_cols <- c('smoking_raw', 'alcohol_raw', 'ethnicity_raw', 'tdi_raw', 'education_raw', 'oily_fish', 'fresh_fruit', 'cooked_vg', 'BMI_raw', 'sex', 'month_birth', 'year_birth', 'diabetes', 'avg_hh_income')


result <- count_exclusions(df_before, full_df)
total_excluded_covd <- result$excluded

# Print total exclusions for step 4.1
print(paste("Total participants excluded in step 4.1:", total_excluded_covd))
```


### 4.2 We remove all rows where any column contains the entry "Prefer not to answer" and
"Don't know":
```{r}


## we transform all the values marked as "Prefer not to answer" and "Don't know" as NAs:
answers_to_remove <- c("Prefer not to answer", "Do not know")

full_df <- full_df %>%
  mutate(across(where(is.character), ~ ifelse(. %in% answers_to_remove, NA, .)))


# we remove the NA amongst the following columns:
df_before <- full_df
keep_cols <- c('smoking', 'alcohol', 'ethnicity', 'tdi', 'education_level', 'oily_fish', 'fresh_fruit', 'cooked_vg', 'BMI', 'red_meat', 'non_oily_fish', 'processed_meat', 'diabetes', 'avg_hh_income')
full_df <- full_df[complete.cases(full_df[, keep_cols]), ]

# we count the exclusions just like we did before
count_exclusions <- function(df_before, df_after) {
  total_before <- nrow(df_before)
  total_after <- nrow(df_after)
  excluded <- total_before - total_after
  return(list(excluded = excluded))
}

result <- count_exclusions(df_before, full_df)
total_excluded_NA <- result$excluded

print(paste("Total participants excluded after removing newly generated NAs:", total_excluded_NA))
```

## Final count:
```{r}
final_n <- nrow(full_df)
print(paste("After applying all exclusions, we end up with a sample of n =", final_n," participants"))
```



############ Inclusion of PA data ############
# Link PA1 and PA2 (Exposure) data

```{r}
library(readr)
enmoTSParallel_PA1 <- list.files(path="/home/hiroshi/projects/UKBB/PA1_parallel_TS_ENMO/outputs/", full.names = TRUE) %>% 
  lapply(read_csv) %>% 
  bind_rows 

ML_derived_PA2 <- fread("/home/hiroshi/projects/UKBB/PA2_category_1020.csv")
ML_derived_PA2 <- ML_derived_PA2 %>%  dplyr::select(
  eid, 
  "40048-0.0",  
  "40049-0.0",
  "40047-0.0", 
  "40046-0.0", 
  "40044-0.0",  
  "40045-0.0",
  "40043-0.0", 
  "40042-0.0" 
  )

```


We create a "age-at-accelerometer-wear variable":

```{r}
# add date of birth
full_df$approx_dob <-
  as.Date(paste(full_df$year_birth, full_df$month_birth, "15", sep = "-"),
          "%Y-%m-%d") # UK Biobank doesn't contain day of birth as it would be unnecessary identifying information, so we roughly impute it as the 15th of the birth month.
# add age at entry in days
full_df$age_entry_days <- as.numeric(difftime(full_df$date_end_accel,
                                              full_df$approx_dob,
                                              units = "days"))
# Convert to age at entry in years
full_df$age_entry_years <- as.double(full_df$age_entry_days)/365.25

# Add age groups
full_df$age_gp <-
  cut(
    full_df$age_entry_years,
    breaks = c(40, 50, 60, 70, 80),
    right = FALSE,
    labels = c("40-49", "50-59", "60-69", "70-79")
  )



boxplot(age_entry_years ~ age_gp, data = full_df)
```

We make a function to cut by quantile:

```{r}
qtile_cut <-  function(x, probs = seq(0, 1, 0.25), na.rm = TRUE, labels = NULL) {
    breaks <- quantile(x = x, probs = probs, na.rm = na.rm)
    out <- cut(x = x, breaks = breaks, labels = labels, right = FALSE, include.lowest = TRUE)
    return(out)
}
```


We cut overall activity and Townsend Deprivation Index into quarters:

```{r}
full_df$overall_activity_quarters <- qtile_cut(full_df$overall_activity, labels = c("Quarter 1", "Quarter 2", "Quarter 3", "Quarter 4"))
# OxWearable Note: The TDI classification here was quarters of the study population, which was used in the example papers. However, our group now typically uses TDI scaled to quarters of the UK population, 
# as listed [here](https://s3-eu-west-1.amazonaws.com/statistics.digitalresources.jisc.ac.uk/dkan/files/Townsend_Deprivation_Scores/UK%20Townsend%20Deprivation%20Scores%20from%202011%20census%20data.pdf, page 15)

full_df$tdi_quarters <- qtile_cut(full_df$tdi, labels = c("Quarter 1", "Quarter 2", "Quarter 3", "Quarter 4"))


boxplot(overall_activity ~ overall_activity_quarters, data = full_df)
boxplot(tdi ~ tdi_quarters, data = full_df)


```




We save the the final df which includes the exlucded participants and the recoding
suggested by OxWearables (as well as the data mapping we had to do)

```{r}
write.csv(full_df, "/home/yacine/UKBB_beluga/df_exc_recode.csv", row.names = FALSE)
```



